---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(123)
library("mlr3torch")
lgr::get_logger("mlr3")$set_threshold("warn")
```



# mlr3torch

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![r-cmd-check](https://github.com/mlr-org/mlr3torch/actions/workflows/r-cmd-check.yml/badge.svg)](https://github.com/mlr-org/mlr3torch/actions/workflows/r-cmd-check.yml)
[![CRAN status](https://www.r-pkg.org/badges/version/mlr3torch)](https://CRAN.R-project.org/package=mlr3torch)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

Deep Learning with torch and mlr3.

## Installation

```{r eval = FALSE}
remotes::install_github("mlr-org/mlr3torch")
```


## Status

`mlr3torch` is currently still unstable but under active development.
Not everything might work yet and the API might change without notice.

## What is mlr3torch?


`mlr3torch` is a deep learning framework for the [`mlr3`](https://mlr-org.com) ecosystem built on top of [`torch`](https://torch.mlverse.org/).
It allows to easily build, train and evaluate deep learning models in a few lines of codes, without needing to worry about low-level details.
Off-the-shelf learners are readily available, but custom architectures can be defined by connection `PipeOpTorch` operators in an `mlr3pipelines::Graph`.

Using predefined learners such as a simple multi layer perceptron (MLP) works just like any other mlr3 `Learner`.

```{r}
library(mlr3torch)
learner_mlp = lrn("classif.mlp",
  # defining network parameters
  activation     = nn_relu,
  layers         = 1,
  d_hidden       = 20,
  # training parameters
  batch_size     = 16,
  epochs         = 50,
  device         = "cpu",
  # Defining the optimizer, loss, and callbacks
  optimizer      = t_opt("adam", lr = 0.1),
  loss           = t_loss("cross_entropy"),
  callbacks      = t_clbk("history"), # this saves the history in the learner
  # Measures to track
  measures_valid = msrs(c("classif.logloss", "classif.ce")),
  measures_train = msrs(c("classif.acc"))
)
```

This learner can for example be used for resampling or benchmarking.

```{r}
resample(
  task       = tsk("iris"),
  learner    = learner_mlp,
  resampling = rsmp("holdout")
)
```

Below, we construct the same architecture using `PipeOpTorch` objects.
The first pipeop -- a `PipeOpTorchIngress` -- defines the entrypoint of the network.
All subsequent pipeops define the neural network layers.

```{r}
architecture = po("torch_ingress_num") %>>%
  po("nn_linear", out_features = 20) %>>%
  po("nn_relu") %>>%
  po("nn_head")
```

To turn this into a learner, we configure the loss, optimizer, callbacks and the training arguments.

```{r}
graph_mlp = architecture %>>%
  po("torch_loss", loss = t_loss("cross_entropy")) %>>%
  po("torch_optimizer", optimizer = t_opt("adam", lr = 0.1)) %>>%
  po("torch_callbacks", callbacks = t_clbk("history")) %>>%
  po("torch_model_classif", batch_size = 16, epochs = 50, device = "cpu")

as_learner(graph_mlp)
```

## Feature Overview

* Off-the-shelf architectures are readily available as `mlr3::Learner`s.
* Custom learners can be defined using the `Graph` language from `mlr3pipelines` or using `nn_module`s.
* The package supports tabular and image data.
* It is possible to customize the training process via (predefined or custom) callbacks.
* The package is fully integrated into the `mlr3` ecosystem.

## Documentation

The easiest way to learn about `mlr3torch` is to read one of the vignettes.

## Acknowledgements

* Without the great R package `torch` none of this would have been possible.
* The names for the callback stages are taken from [luz](https://mlverse.github.io/luz/), another high-level deep learning framework for R `torch`.
* Building neural networks using `PipeOpTorch` operators is inspired by [keras](https://keras.io/).

## Bugs, Questions, Feedback

*mlr3torch* is a free and open source software project that
encourages participation and feedback. If you have any issues,
questions, suggestions or feedback, please do not hesitate to open an
“issue” about it on the GitHub page\!

In case of problems / bugs, it is often helpful if you provide a
“minimum working example” that showcases the behaviour (but don’t
worry about this if the bug is obvious).

Please understand that the resources of the project are limited:
response may sometimes be delayed by a few days, and some feature
suggestions may be rejected if they are deemed too tangential to the
vision behind the project.
