---
title: Data Types in `mlr3torch`
output: html_document
---


Before we start covering the image representations of `mlr3torch`, we start with a first primer on data representations in `mlr3`.
The data in `mlr3` is accessed through a `DataBackend`.
The most common backend is the `DataBackendDataTable`.
To create such a backend, `as_data_backend()` can be used as shown below.

```{r}
library(mlr3verse)
library(data.table)

backend_dt = as_data_backend(mtcars)
backend_dt
```

The most important interface to such a dataset is that row- and columnwise subsets of the data can be queried using the `$data()` method.

```{r}
backend_dt$data(1:2, c("mpg", "cyl"))
```

When focusing on multi-modal data types in mlr3torch -- e.g. datasets with images and tabular features -- it is desireable to be able to represent images alongside tabular data.
For this reason, `mlr3torch` provides two data types for images, namely:
* `imageuri` - Essentially a `character()` vector containing paths to images on disk.

The former can be used when the dataset is small and fits into RAM (e.g. MNIST), while the latter is intended to be used for larger datasets.


A central assumption in `mlr3` is that data is represented in *some* row-column format.
Howver, differ

```{r}
library(mlr3torch)

task = tsk("mtcars")
task$data()
```



