---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Quickstart

In this vignette we will show how to get started with `mlr3torch` by training a simple neural network on a tabular regression problem.
We assume that you are familiar with the `mlr3` framework, see e.g. the [mlr3 book](https://mlr3book.mlr-org.com/).
As a first example, we will train a simple multi-layer perceptron (MLP) on the well-known "mtcars" task, where the goal is to predict the miles per galleon ('mpg') of cars.
This architecture comes as a predfined learner with `mlr3torch`, but you can also easily create new network architectures, see the *Neural Networks as Graphs* vignette for a detailed introduoduion.
We first set a seed for reproducibility, load the library and construct the task.

```{r, message = FALSE}
set.seed(314)
library(mlr3torch)
task = tsk("mtcars")
task$head()
```

Learners in `mlr3torch` work very similary to other `mlr3` learners.
Below, we construct a simple multi layer perceptron for regression.
We do this as usual by calling `lrn()` and configuring the parameters:
We use two hidden layers with 50 neurons,
For training, we set the batch size to 32, the number of training epochs to 30 and the device to `"cpu"`.
For a complete description of the available parameters see `?mlr3torch::LearnerTorchMLP`.

```{r}
mlp = lrn("regr.mlp",
  # architecture parameters
  neurons = c(50, 50),
  # training arguments
  batch_size = 32, epochs = 30, device = "cpu"
)
mlp
```

We can use this learner for training and prediction just like any other regression learner.
Below, we split the observations into a training and test set, train the learner on the training set and create predictions for the test set.
Finally, we compute the mean squared error of the predictions.

```{r}
# Split the obersevations into training and test set
splits = partition(task)
# Train the learner on the train set
mlp$train(task, row_ids = splits$train)
# Predict the test set
prediction = mlp$predict(task, row_ids = splits$test)
# Compute the mse
prediction$score(msr("regr.mse"))
```

## Configuring a Learner

Although torch learners are quite like other `mlr3` learners, there are some differences.
One is that all `LearnerTorch` classes have *construction arguments*, i.e. torch learners are more modular than other learners.
While learners are free to implement their own construction arguments, there are some that are common to all torch learners, namely the `loss`, `optimizer` and `callbacks`.
Each of these object can have their own parameters that are included in the `LearnerTorch`'s parameter set.

In the previous example, we did not specify any of these explicitly and used the default values, which was the Adam optimizer, MSE as the loss and no callbacks.
We will now show how to configure these three aspects of a learner through the `mlr3torch::TorchOptimizer`, `mlr3torch::TorchLoss`, and `mlr3torch::TorchCallback` classes.

### Loss

The loss function, also known as the objective function or cost function, measures the discrepancy between the predicted output and the true output. It quantifies how well the model is performing during training.
The R package `torch`, which underpins the `mlr3torch` framework, already provides a number of predefined loss functions such as the Mean Squared Error (`nn_mse_loss`), the Mean Absolute Error (`nn_l1_loss`), or the cross entropy loss (`nn_cross_entropy_loss`).
In `mlr3torch`, we represent loss functions using the `mlr3torch::TorchLoss` class.
It provides a thin wrapper around the torch loss functions and annotates them with meta information, most importantly a `paradox::ParamSet` that allows to configure the loss function.
Such an object can be constructed using `t_loss(<key>)`.
Below, we construct the L1 loss function, which is also known as Mean Absolute Error (MAE).
The printed output below informs us about the wrapped loss function `(nn_l1_loss`), the configured parameters, the packages it depends on and for which task types it can be used.

```{r}
l1 = t_loss("l1")
l1
```

Its `ParamSet` contains only one parameter, namely `reduction`, which specifies how the loss is reduced over the batch.

```{r}
# the paradox::ParamSet of the loss
l1$param_set
```

The wrapped loss module generator is accessible through the slot `$generator`.

```{r}
l1$generator
```

We can pass the `TorchLoss` as the argument `loss` during initialization of the learner.
The parameters of the loss are added to the learner's `ParamSet`, prefixed with `"loss."`.

```{r}
mlp_l1 = lrn("regr.mlp", loss = l1)
mlp_l1$param_set$values$loss.reduction
```

All predefined loss functions are stored in the `mlr3torch_losses` dictionary, from which they can be retrieved using `t_loss(<key>)`.

```{r}
mlr3torch_losses
```

### Optimizer

The optimizer determines how the model's weights are updated based on the calculated loss. It adjusts the parameters of the model to minimize the loss function, optimizing the model's performance.
Optimizers work analogous to loss functions, i.e. `mlr3torch` provides a thin wrapper -- the `TorchOptimizer` class -- around the optimizers such as Adam (`optim_adam`) or SGD (`optim_sgd`).
`TorchLoss` objects can be constructed using `t_opt(<key>)`.
For optimizers, the associated `ParamSet` is more interesting as we see below:

```{r}
sgd = t_opt("sgd")
sgd

sgd$param_set
```

The wrapped torch optimizer can be accessed through the slot `generator`.

Parameters of `TorchOptimizer` (but also `TorchLoss` and `TorchCallback`) can be set in the usual `mlr3` way, i.e. either during construction, or afterwards using the `$set_values()` method of the parameter set.

```{r}
sgd$param_set$set_values(
  lr = 0.5, # increase learning rate
  nesterov = FALSE # no nesterov momentum
)
```

Below we see that the optimizer's parameters are added to the learner's `ParamSet` (prefixed with `"opt."`) and that the values are set to the values we specified.

```{r}
mlp_sgd = lrn("regr.mlp", optimizer = sgd)
as.data.table(mlp_sgd$param_set)[
  startsWith(id, "opt.")][[1L]]
mlp_sgd$param_set$values[c("opt.lr", "opt.nesterov")]
```

By exposing the optimizer's parameters, they can be conveniently tuned using [`mlr3tuning`](https://github.com/mlr-org/mlr3tuning).

All available optimizers are stored in the `mlr3torch_optimizers` dictionary.

```{r}
mlr3torch_optimizers
```

### Callbacks

The third important configuration option are callbacks which allow to customize the training process.
This allows saving model checkpoints, logging metrics, or implementing custom functionality for specific training scenarios.
For a tutorial on how to implement a custom callback, see the *Custom Callbacks* vignette.
Here, we will only show how to use predefined callbacks.
Below, we retrieve the `"history"` callback using `t_clbk()`, which has no parameters and merely saves the training and validation history in the learner so it can be accessed afterwards.

```{r}
history = t_clbk("history")
history
```

If we wanted to learn about what the callback does, we can access the help page of the wrapped object using the `$help()` method.
Note that this is also possible for the loss and optimizer.

```{r, eval = FALSE}
history$help()
```

All predefined callbacks are stored in the `mlr3torch_callbacks` dictionary.

```{r}
mlr3torch_callbacks
```

### Putting it Together

We now define our customized MLP learner using the loss, optimizer and callback we have just covered.
To keep track of the performance, we use 30% of the training data for validation and evaluate it using the MAE `Measure`.
Note that the `mearures_valid` and `measures_train` parameters of `LearnerTorch` take common `mlr3::Measure`s, whereas the loss function must be a `TorchLoss`.

```{r}
mlp_custom = lrn("regr.mlp",
  # construction arguments
  optimizer = sgd, loss = l1, callbacks = history,
  # scores to keep track of
  measures_valid = msr("regr.mae"),
  # other parameters are left as-is:
  # architecture
  neurons = c(50, 50),
  # training arguments
  batch_size = 32, epochs = 30, device = "cpu",
  # validation proportion
  validate = 0.3
)

mlp_custom
```

We now train the learner on the "mtcars" task again and use the same train-test split as before.

```{r}
mlp_custom$train(task, row_ids = splits$train)
prediction_custom = mlp_custom$predict(task, row_ids = splits$test)
```

Below we make predictions on the unseen test data and compare the scores.
Because we directly optimized the L1 (aka MAE) loss and tweaked the learning rate, our configured `mlp_custom` learner has a lower MAE than the default `mlp` learner.

```{r}
prediction_custom$score(msr("regr.mae"))
prediction$score(msr("regr.mae"))
```

Because we configured the learner to use the history callback, we can find the validation history in its `$model` slot:

```{r}
head(mlp_custom$model$callbacks$history$valid)
```

The plot below shows it for the epochs 6 to 30.

```{r, fig.align='center', out.width="50%", fig.dpi = 300, echo = FALSE}
mlp_custom$param_set$set_values(
  measures_valid = msr("regr.mae")
)

data = data.table::melt(mlp_custom$model$callbacks$history$valid,
  id.vars = "epoch", variable.name = "measure", value.name = "score")
ggplot2::ggplot(data = data[6:30, ], ggplot2::aes(x = epoch, y = score)) +
  viridis::scale_color_viridis(discrete = TRUE) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::labs(
    x = "Epoch",
    y = "MAE",
    title = "Validation Loss"
  ) +
  ggplot2::theme_bw()
```

Other important information that is stored in the `Learner`'s model is the `$network`, which is the underlying `nn_module`.
For a full description of the model, see `?LearnerTorch`.
