---
title: "Adding a Custom Task Type"
author: "Sebastian Fischer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Adding a Custom Task Type}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
PARKED, NOT BUILT.

This is the "do it properly" counterpart to vignettes/articles/custom_task_type.Rmd, which covers
only the quick route via `TaskTorch`. It walks through adding a real task type to mlr3, using
multi-label classification and a variational autoencoder as examples.

It rendered cleanly before it was displaced. It has NOT been re-rendered since being parked here,
so re-render it before promoting it back to vignettes/articles/. See TODOs.md.
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lgr::get_logger("mlr3")$set_threshold("warn")
set.seed(314)
```

`mlr3torch` supports classification (`"classif"`) and regression (`"regr"`) out of the box.
Not every learning problem fits into one of these two, though.
Multi-label classification attaches several binary labels to each observation, survival analysis predicts a time and a censoring status, and an autoencoder has no target column at all.
In `mlr3`, each of these is a separate *task type*, and this vignette shows how to add one and how to make `mlr3torch` learners work with it.

We will do this twice.
The first example is multi-label classification, which is supervised and therefore follows the same pattern as the built-in task types.
The second example is a variational autoencoder, which is unsupervised and shows what changes when a task has no target.

The work always splits into two halves.
The first half registers the task type with `mlr3` and is not specific to deep learning at all -- you would write exactly the same code to support the task type in `mlr3learners`.
The second half consists of a small number of S3 methods that tell `mlr3torch` how to move between the world of `mlr3` objects and the world of tensors.

If you only want to run a custom problem through `mlr3torch` and do not intend to hand the task type to anyone else, you do not have to write any of this: see the *Custom Learning Problems* article and `?TaskTorch` for the quick route.

```{r, message = FALSE}
library(mlr3torch)
library(mlr3misc)
library(data.table)
library(checkmate)
library(R6)
```

## The Ingredients of a Task Type {#sec-ingredients}

Before we start writing code, it is worth listing what a task type actually consists of.
On the `mlr3` side you need:

* Entries in `mlr_reflections`, the central registry that says which task types exist and which classes implement them.
* A `Task` class that holds the data and knows which columns are the target.
* A `Prediction` class and the `PredictionData` S3 methods that `mlr3` uses to combine, filter and validate what a learner returns.
* At least one `Measure`, so that predictions can be scored.

On the `mlr3torch` side you need to answer three questions, each of which is an S3 generic that dispatches on the task:

* `output_dim_for()`: how many units does the output layer of the network need?
* `get_target_batchgetter()`: how is the target of a batch turned into a tensor?
* `encode_prediction()`: how is the output of the network turned back into a prediction?

Finally, a `TorchLoss` has to declare that it is applicable to the new task type, because `mlr3torch` only knows a sensible default loss for classification and regression.

## Multi-Label Classification {#sec-multilabel}

In multi-label classification each observation carries several binary labels at once, and the labels are not mutually exclusive.
A weather observation can be sunny and warm and windy, or sunny and warm but not windy.
We represent such a task as a table with one `logical` column per label.

### Registering the Task Type {#sec-multilabel-reflections}

`mlr_reflections` is a central registry that `mlr3` consults whenever it needs to know something about a task type.
Adding a row to `mlr_reflections$task_types` announces the new type and names the classes that implement it:

```{r}
mlr_reflections$task_types = setkeyv(rbind(mlr_reflections$task_types, rowwise_table(
  ~type, ~package, ~task, ~learner, ~prediction, ~prediction_data, ~measure,
  "multilabel", "mlr3torch", "TaskMultilabel", "LearnerMultilabel", "PredictionMultilabel", "PredictionDataMultilabel", "MeasureMultilabel"
), fill = TRUE), "type")

mlr_reflections$task_types["multilabel", ]
```

The classes are looked up by name, so the strings have to match the classes we define below.
The `task` and `prediction` entries must name classes that really exist, because other packages rely on them.
`mlr3pipelines`, for example, uses them as the input and output channel types of a `PipeOpLearner`.
The `learner` and `measure` entries name the *base* class of the type and are only consulted when `mlr3` checks whether two related task types are compatible.
We follow the naming convention here, but our learner will inherit from `LearnerTorch` (which inherits from `Learner` directly) rather than from an intermediate `LearnerMultilabel` class.

Several further registries are indexed by task type.
Column roles, task properties, learner properties and measure properties behave exactly as they do for regression, so we simply copy those entries:

```{r}
mlr_reflections$task_col_roles$multilabel = mlr_reflections$task_col_roles$regr
mlr_reflections$task_properties$multilabel = mlr_reflections$task_properties$regr
mlr_reflections$learner_properties$multilabel = mlr_reflections$learner_properties$regr
mlr_reflections$measure_properties$multilabel = mlr_reflections$measure_properties$regr
```

The predict types, on the other hand, are our own design decision.
A multi-label learner can predict the labels themselves (`"response"`) or one probability per label (`"prob"`).
The list maps each predict type to everything it implies, so a learner with `predict_type = "prob"` also produces a `"response"`:

```{r}
mlr_reflections$learner_predict_types$multilabel = list(
  response = "response",
  prob = c("response", "prob")
)
```

Finally, `mlr3` needs a default measure for the task type, which is used whenever `$score()` or `$aggregate()` is called without arguments.
We register the id here and define the measure itself in the section on a custom measure.

```{r}
mlr_reflections$default_measures$multilabel = "multilabel.hamming"
```

Note that the code above mutates the state of `mlr3`.
In a package, do this inside an `.onLoad()` hook rather than at the top level of an R file, because top-level code runs when the package is *built*, not when it is loaded.

### The Task {#sec-multilabel-task}

The `Task` class is the easy part.
`mlr3::Task` already supports more than one target column, so all we have to do is pass our `task_type` to the parent class and check that the target columns are the sort of columns we expect:

```{r}
TaskMultilabel = R6Class("TaskMultilabel",
  inherit = TaskSupervised,
  public = list(
    initialize = function(id, backend, target, label = NA_character_) {
      assert_character(target, min.len = 1L, any.missing = FALSE)
      super$initialize(id = id, task_type = "multilabel", backend = backend,
        target = target, label = label)
      types = self$col_info[list(target), "type", on = "id"][[1L]]
      if (!all(types == "logical")) {
        stopf("All targets of a multilabel task must be logical.")
      }
    }
  )
)
```

`Task$truth()` returns `self$data(rows, cols = self$target_names)`, which for us is a `data.table` with one logical column per label.
That is exactly what we want, so we do not need to override it.

Next we need some data to work with.
The three labels below depend on the five features in different ways, and `warm` depends on an *interaction* of two features, which a linear model could not pick up:

```{r}
n = 500
dat = data.table(x1 = rnorm(n), x2 = rnorm(n), x3 = rnorm(n), x4 = rnorm(n), x5 = rnorm(n))
dat[, `:=`(
  sunny = x1 + x2 > 0,
  warm = x2 * x3 > 0,
  windy = x4 - x5 > 0
)]

tsk_weather = TaskMultilabel$new("weather", dat, target = c("sunny", "warm", "windy"))
tsk_weather
```

The printed output shows the three targets and the five features, just as it would for any other task type.

### The Prediction {#sec-multilabel-prediction}

A `Prediction` in `mlr3` is a thin wrapper around a *prediction data* object.
The latter is a plain `list()` with a class attribute, holding the `row_ids`, the `truth`, and one element per predict type.
The class name of that list is the one we registered in the `prediction_data` column above.

```{r}
PredictionMultilabel = R6Class("PredictionMultilabel",
  inherit = Prediction,
  public = list(
    initialize = function(task = NULL, row_ids = task$row_ids, truth = task$truth(),
      response = NULL, prob = NULL, check = TRUE) {
      pdata = discard(list(row_ids = row_ids, truth = truth, response = response, prob = prob), is.null)
      class(pdata) = c("PredictionDataMultilabel", "PredictionData")
      if (check) pdata = check_prediction_data(pdata)

      self$task_type = "multilabel"
      self$man = "mlr3torch::PredictionMultilabel"
      self$data = pdata
      self$predict_types = intersect(c("response", "prob"), names(pdata))
    }
  ),
  active = list(
    # a data.table() with one logical column per label
    response = function(rhs) {
      if (!missing(rhs)) stopf("Field is read-only")
      self$data$response
    },
    # a matrix() with one numeric column per label
    prob = function(rhs) {
      if (!missing(rhs)) stopf("Field is read-only")
      self$data$prob
    }
  )
)
```

The `Prediction` base class already provides `$row_ids`, `$truth` and `$missing`, so we only add active bindings for our own predict types.

### Prediction Data {#sec-multilabel-pdata}

Prediction data is what learners actually return and what `mlr3` passes around internally, because it is cheaper to combine and to subset than a full `Prediction` object.
`mlr3` interacts with it through a small set of S3 generics, and each of them needs a method for our `PredictionDataMultilabel` class.

`check_prediction_data()` validates and normalizes what a learner returned.
This is the right place to bring the pieces into a canonical form.
Below we make sure that the label order matches the truth, and we derive the `response` from the probabilities if a learner only supplies the latter:

```{r}
check_prediction_data.PredictionDataMultilabel = function(pdata, ...) {
  n = length(assert_row_ids(pdata$row_ids))
  labels = names(pdata$truth)
  pdata$truth = assert_data_table(pdata$truth, nrows = n, types = "logical")

  if (!is.null(pdata$response)) {
    pdata$response = assert_data_table(as.data.table(pdata$response), nrows = n, types = "logical")
    assert_names(names(pdata$response), permutation.of = labels)
    setcolorder(pdata$response, labels)
  }
  if (!is.null(pdata$prob)) {
    assert_matrix(pdata$prob, nrows = n, ncols = length(labels), mode = "numeric")
    assert_names(colnames(pdata$prob), permutation.of = labels)
    pdata$prob = pdata$prob[, labels, drop = FALSE]
    if (is.null(pdata$response)) {
      pdata$response = as.data.table(pdata$prob > 0.5)
    }
  }
  pdata
}
```

`is_missing_prediction_data()` reports for which rows the learner failed to predict, which `mlr3` uses to warn about incomplete predictions:

```{r}
is_missing_prediction_data.PredictionDataMultilabel = function(pdata, ...) {
  miss = logical(length(pdata$row_ids))
  if (!is.null(pdata$response)) {
    miss = Reduce(`|`, lapply(pdata$response, is.na))
  }
  pdata$row_ids[miss]
}
```

`as_prediction()` turns prediction data back into the `Prediction` object:

```{r}
as_prediction.PredictionDataMultilabel = function(x, check = TRUE, ...) {
  invoke(PredictionMultilabel$new, check = check, .args = x)
}
```

`c()` combines the predictions of the individual resampling iterations into a single one and `filter_prediction_data()` subsets them.
Without these two, `resample()` and `benchmark()` do not work:

```{r}
c.PredictionDataMultilabel = function(..., keep_duplicates = TRUE) {
  dots = list(...)
  assert_list(dots, "PredictionDataMultilabel")
  if (length(dots) == 1L) return(dots[[1L]])

  row_ids = do.call(c, map(dots, "row_ids"))
  truth = rbindlist(map(dots, "truth"))
  response = if (!is.null(dots[[1L]]$response)) rbindlist(map(dots, "response"))
  prob = if (!is.null(dots[[1L]]$prob)) do.call(rbind, map(dots, "prob"))

  if (!keep_duplicates) {
    keep = !duplicated(row_ids, fromLast = TRUE)
    row_ids = row_ids[keep]
    truth = truth[keep]
    if (!is.null(response)) response = response[keep]
    if (!is.null(prob)) prob = prob[keep, , drop = FALSE]
  }

  pdata = discard(list(row_ids = row_ids, truth = truth, response = response, prob = prob), is.null)
  class(pdata) = c("PredictionDataMultilabel", "PredictionData")
  pdata
}

filter_prediction_data.PredictionDataMultilabel = function(pdata, row_ids, ...) {
  keep = pdata$row_ids %in% row_ids
  pdata$row_ids = pdata$row_ids[keep]
  pdata$truth = pdata$truth[keep]
  if (!is.null(pdata$response)) pdata$response = pdata$response[keep]
  if (!is.null(pdata$prob)) pdata$prob = pdata$prob[keep, , drop = FALSE]
  pdata
}
```

Lastly, `Prediction$print()` displays `as.data.table(prediction)`, so we flatten the two tables into prefixed columns.
Note the `copy()` calls: `setnames()` renames by reference, and without them printing a prediction would rename the columns that the prediction itself holds.

```{r}
as.data.table.PredictionMultilabel = function(x, ...) {
  tab = data.table(row_ids = x$data$row_ids)
  tab = cbind(tab, setnames(copy(x$data$truth), paste0("truth.", names(x$data$truth))))
  if (!is.null(x$data$response)) {
    tab = cbind(tab, setnames(copy(x$data$response), paste0("response.", names(x$data$response))))
  }
  tab
}
```

That is the whole `mlr3` side.
We can now build a prediction by hand, without any learner being involved.
The one below always guesses `TRUE`:

```{r}
prediction = PredictionMultilabel$new(
  task = tsk_weather,
  response = as.data.table(matrix(TRUE, nrow = tsk_weather$nrow, ncol = 3,
    dimnames = list(NULL, tsk_weather$target_names)))
)
prediction
```

### A Custom Measure {#sec-multilabel-measure}

A `Measure` scores a `Prediction`.
Its `task_type` ties it to our new task type, `predict_type` declares what it needs from the learner, and the private `$.score()` method does the actual work.
The classic multi-label metric is the *Hamming loss*, the fraction of individual label decisions that are wrong:

```{r}
MeasureMultilabelHamming = R6Class("MeasureMultilabelHamming",
  inherit = Measure,
  public = list(
    initialize = function() {
      super$initialize(
        id = "multilabel.hamming",
        task_type = "multilabel",
        predict_type = "response",
        range = c(0, 1),
        minimize = TRUE,
        label = "Hamming Loss"
      )
    }
  ),
  private = list(
    .score = function(prediction, ...) {
      mean(as.matrix(prediction$truth) != as.matrix(prediction$response))
    }
  )
)

mlr_measures$add("multilabel.hamming", MeasureMultilabelHamming)
```

Adding the measure to the `mlr_measures` dictionary makes it available through `msr()`, and we can score the constant prediction from above:

```{r}
prediction$score(msr("multilabel.hamming"))
```

Roughly half of all label decisions are wrong, which is what we expect from a learner that ignores the features.

A measure that consumes probabilities works the same way; it only sets `predict_type = "prob"` and reads `prediction$prob` instead.
The mean binary log-loss over all labels is one example:

```{r}
MeasureMultilabelLogloss = R6Class("MeasureMultilabelLogloss",
  inherit = Measure,
  public = list(
    initialize = function() {
      super$initialize(
        id = "multilabel.logloss",
        task_type = "multilabel",
        predict_type = "prob",
        range = c(0, Inf),
        minimize = TRUE,
        label = "Mean Binary Log-Loss"
      )
    }
  ),
  private = list(
    .score = function(prediction, ...) {
      eps = 1e-7
      truth = as.matrix(prediction$truth)
      prob = pmin(pmax(prediction$prob, eps), 1 - eps)
      -mean(truth * log(prob) + (1 - truth) * log(1 - prob))
    }
  )
)

mlr_measures$add("multilabel.logloss", MeasureMultilabelLogloss)
```

### Teaching mlr3torch about the Task Type {#sec-multilabel-torch}

We now switch to the second half, which is where `mlr3torch` comes in.
`output_dim_for()` states how many units the output layer of the network needs.
For multi-label classification we predict one logit per label:

```{r}
output_dim_for.TaskMultilabel = function(x, ...) {
  length(x$target_names)
}
```

Helpers such as `po("nn_head")` call this generic, so a network that is built with the `nn()` graph language automatically gets the right output size.

`get_target_batchgetter()` returns a function that converts the target columns of a batch -- a `data.table` -- into the tensor `y` that the loss is applied to.
For us that is a `(batch_size, n_labels)` float tensor of zeros and ones, which is what `nn_bce_with_logits_loss()` expects:

```{r}
get_target_batchgetter.TaskMultilabel = function(task, ...) {
  function(data) {
    torch_tensor(as.matrix(data), dtype = torch_float())
  }
}
```

`encode_prediction()` is the inverse of the target encoding.
It maps the raw output of the network to a `list()` that `mlr3::as_prediction_data()` understands, i.e. a list whose elements are named after our predict types.
Our network emits logits, so we apply a sigmoid per label and threshold at `0.5`:

```{r}
encode_prediction.TaskMultilabel = function(task, predict_tensor, predict_type, ...) {
  prob = as.matrix(with_no_grad(torch_sigmoid(predict_tensor))$cpu())
  colnames(prob) = task$target_names
  response = as.data.table(prob > 0.5)

  if (predict_type == "prob") {
    list(response = response, prob = prob)
  } else {
    list(response = response)
  }
}
```

An individual learner can always deviate from this by overwriting its private `$.encode_prediction()` method.
Implementing the generic is what makes the *generic* learners, such as `LearnerTorchModel`, work for the new task type.

The last ingredient is a loss function.
A `TorchLoss` declares which task types it is applicable to, so we wrap `torch::nn_bce_with_logits_loss()` and register it in the `mlr3torch_losses` dictionary:

```{r}
mlr3torch_losses$add("multilabel_bce", TorchLoss$new(
  torch_loss = nn_bce_with_logits_loss,
  task_types = "multilabel",
  id = "multilabel_bce",
  label = "Binary Cross Entropy with Logits"
))

t_loss("multilabel_bce")
```

`mlr3torch` only knows a default loss for classification and regression.
Learners for a custom task type therefore have to supply one, either as a default in their constructor, as we do below, or by leaving the choice to the user.

### A Learner {#sec-multilabel-learner}

At this point writing the learner is no different from writing one for classification or regression, which is described in the documentation of `LearnerTorch`.
We pass `task_type = "multilabel"` and implement the private `$.network()` and `$.ingress_tokens()` methods:

```{r}
LearnerMultilabelMLP = R6Class("LearnerMultilabelMLP",
  inherit = LearnerTorch,
  public = list(
    initialize = function(optimizer = NULL, loss = t_loss("multilabel_bce"), callbacks = list()) {
      param_set = ps(
        neurons = p_uty(tags = "train", custom_check = crate(function(x) checkmate::check_integerish(x, any.missing = FALSE)))
      )
      param_set$set_values(neurons = 20L)

      super$initialize(
        id = "multilabel.mlp",
        task_type = "multilabel",
        label = "Multilabel Multi Layer Perceptron",
        param_set = param_set,
        optimizer = optimizer,
        loss = loss,
        callbacks = callbacks,
        feature_types = c("numeric", "integer"),
        man = "mlr3torch::mlr_learners_multilabel.mlp"
      )
    }
  ),
  private = list(
    .network = function(task, param_vals) {
      layers = list()
      prev = length(task$feature_names)
      for (n in param_vals$neurons) {
        layers = c(layers, list(nn_linear(prev, n), nn_relu()))
        prev = n
      }
      invoke(nn_sequential, .args = c(layers, list(nn_linear(prev, output_dim_for(task)))))
    },
    .ingress_tokens = function(task, param_vals) {
      list(input = ingress_num(shape = c(NA, length(task$feature_names))))
    }
  )
)
```

Because we registered `"prob"` as a predict type and `encode_prediction.TaskMultilabel()` produces it, the learner supports both predict types without any further work:

```{r}
lrn_multilabel = LearnerMultilabelMLP$new()
lrn_multilabel$predict_types
```

### Training and Predicting {#sec-multilabel-train}

The learner now behaves like any other `mlr3` learner.
We configure it, split the task into a training and a test set with `partition()`, and train on the training set:

```{r}
lrn_multilabel$param_set$set_values(
  epochs = 50, batch_size = 32, neurons = c(20, 20), opt.lr = 0.01
)
lrn_multilabel$predict_type = "prob"

splits = partition(tsk_weather)
lrn_multilabel$train(tsk_weather, row_ids = splits$train)
prediction = lrn_multilabel$predict(tsk_weather, row_ids = splits$test)
prediction
```

The `truth.*` columns are the ground truth extracted from the task and the `response.*` columns are what the network predicted.
Both come from the `as.data.table()` method that we wrote in the section on prediction data.
Scoring works with either of our two measures:

```{r}
prediction$score(msrs(c("multilabel.hamming", "multilabel.logloss")))
```

Resampling works as well, which is what the `c()` and `filter_prediction_data()` methods were needed for:

```{r}
rr = resample(tsk_weather, lrn_multilabel, rsmp("cv", folds = 3))
rr$aggregate(msr("multilabel.hamming"))
```

Our custom measure can also be used for validation and early stopping, exactly like a built-in one.
Below we hold out 30% of the training data, score it after every epoch, and stop once the Hamming loss has not improved for five epochs:

```{r}
lrn_early_stopping = LearnerMultilabelMLP$new()
lrn_early_stopping$param_set$set_values(
  epochs = 100, batch_size = 32, neurons = c(20, 20), opt.lr = 0.01,
  patience = 5, measures_valid = msr("multilabel.hamming")
)
lrn_early_stopping$validate = 0.3

lrn_early_stopping$train(tsk_weather)
lrn_early_stopping$internal_valid_scores
lrn_early_stopping$internal_tuned_values
```

Training stopped well before the configured 100 epochs, and `$internal_tuned_values` reports the epoch that `mlr3tuning` would use for internal tuning.

### Using the Graph Language {#sec-multilabel-graph}

Everything said in the [Defining an Architecture](pipeop_torch.html) vignette applies to the new task type as well.
The only difference is that we construct the terminal `PipeOpTorchModel` for our task type ourselves, instead of using the `po("torch_model_classif")` and `po("torch_model_regr")` shorthands:

```{r}
architecture = po("torch_ingress_num") %>>%
  nn("linear_1", out_features = 20) %>>%
  nn("relu_1") %>>%
  nn("linear_2", out_features = 20) %>>%
  nn("relu_2") %>>%
  nn("head") %>>%
  po("torch_loss", t_loss("multilabel_bce")) %>>%
  po("torch_optimizer", "adam", lr = 0.01) %>>%
  PipeOpTorchModel$new("multilabel", param_vals = list(batch_size = 32, epochs = 50))

glrn_multilabel = as_learner(architecture)
glrn_multilabel$train(tsk_weather, row_ids = splits$train)
glrn_multilabel$predict(tsk_weather, row_ids = splits$test)$score(msr("multilabel.hamming"))
```

Note that `nn("head")` sized its output layer through our `output_dim_for()` method, and that the resulting `LearnerTorchModel` encoded its predictions through our `encode_prediction()` method.
Neither of them knows anything about multi-label classification.

If you want a `po("torch_model_multilabel")` shorthand, add a thin subclass to the `mlr_pipeops` dictionary, in the same way that `mlr3torch` does it for classification and regression:

```{r}
PipeOpTorchModelMultilabel = R6Class("PipeOpTorchModelMultilabel",
  inherit = PipeOpTorchModel,
  public = list(
    initialize = function(id = "torch_model_multilabel", param_vals = list()) {
      super$initialize(task_type = "multilabel", id = id, param_vals = param_vals)
    }
  )
)

mlr_pipeops$add("torch_model_multilabel", PipeOpTorchModelMultilabel)
po("torch_model_multilabel", epochs = 50, batch_size = 32)
```

The input and output channels of the new `PipeOp` are typed with the `TaskMultilabel` and `PredictionMultilabel` classes that we registered when registering the task type.

## An Unsupervised Task Type {#sec-vae}

Multi-label classification is supervised, so most of the work above was mechanical.
The second example is more interesting, because it drops the assumption that a task has a target at all.

A variational autoencoder (VAE) compresses each observation into a low-dimensional latent code and reconstructs the input from it.
There is nothing to predict from the outside: the input *is* the target.
This changes three things compared to multi-label classification, and we will look at each of them in turn:

* The task inherits from `TaskUnsupervised`, so its predictions have no `truth` and the measure has to read the ground truth from the task instead.
* The network returns three tensors during training -- the reconstruction, the latent mean and the latent log-variance -- because the loss needs all three.
* The dataset has to produce a target that is not stored in any target column, which is where the private `$.dataset()` method comes in.

### Registering the Task Type {#sec-vae-reflections}

The registration follows exactly the same pattern as before, except that the column roles and task properties are copied from `"unsupervised"` rather than from `"regr"`:

```{r}
mlr_reflections$task_types = setkeyv(rbind(mlr_reflections$task_types, rowwise_table(
  ~type, ~package, ~task, ~learner, ~prediction, ~prediction_data, ~measure,
  "vae", "mlr3torch", "TaskVAE", "LearnerVAE", "PredictionVAE", "PredictionDataVAE", "MeasureVAE"
), fill = TRUE), "type")

mlr_reflections$task_col_roles$vae = mlr_reflections$task_col_roles$unsupervised
mlr_reflections$task_properties$vae = mlr_reflections$task_properties$unsupervised
mlr_reflections$learner_properties$vae = mlr_reflections$learner_properties$regr
mlr_reflections$measure_properties$vae = mlr_reflections$measure_properties$regr
mlr_reflections$learner_predict_types$vae = list(response = "response")
mlr_reflections$default_measures$vae = "vae.reconstruction"
```

There is only one predict type, `"response"`, which will hold the reconstruction of the input.

Note that `mlr3` already registers a task type `"unsupervised"`, but its `prediction`, `prediction_data` and `measure` entries are all `NA` and it has no predict types.
It exists so that `PipeOp`s can pass target-less data around, not so that learners can train on it -- building a learner for it fails with *"Assertion on 'predict_types' failed: Must be a subset of {}"*.
A task type that is going to be predicted on therefore has to be registered separately, as we do here.

### The Task and the Prediction {#sec-vae-task}

`TaskUnsupervised` is the base class for tasks without a target, so the task class is a one-liner:

```{r}
TaskVAE = R6Class("TaskVAE",
  inherit = TaskUnsupervised,
  public = list(
    initialize = function(id, backend, label = NA_character_) {
      super$initialize(id = id, task_type = "vae", backend = backend, label = label)
    }
  )
)
```

As data we use the four numeric features of the iris dataset, standardized so that a reconstruction error is easy to interpret:

```{r}
iris_scaled = as.data.table(scale(iris[, 1:4]))
tsk_iris_unsup = TaskVAE$new("iris_unsup", iris_scaled)
tsk_iris_unsup
```

The printed output confirms that the task has four features and no target.

The prediction class is the same shape as `PredictionMultilabel`, but simpler.
There is no `truth` -- `mlr3` only extracts a truth from tasks that inherit from `TaskSupervised` -- and the single predict type is a numeric matrix with one column per feature:

```{r}
PredictionVAE = R6Class("PredictionVAE",
  inherit = Prediction,
  public = list(
    initialize = function(task = NULL, row_ids = task$row_ids, response = NULL, check = TRUE) {
      pdata = discard(list(row_ids = row_ids, response = response), is.null)
      class(pdata) = c("PredictionDataVAE", "PredictionData")
      if (check) pdata = check_prediction_data(pdata)

      self$task_type = "vae"
      self$man = "mlr3torch::PredictionVAE"
      self$data = pdata
      self$predict_types = "response"
    }
  ),
  active = list(
    # a matrix() with one numeric column per feature
    response = function(rhs) {
      if (!missing(rhs)) stopf("Field is read-only")
      self$data$response
    }
  )
)
```

The prediction data methods are the same five as for multi-label classification, only shorter, because there is a single matrix to look after:

```{r}
check_prediction_data.PredictionDataVAE = function(pdata, ...) {
  n = length(assert_row_ids(pdata$row_ids))
  assert_matrix(pdata$response, nrows = n, mode = "numeric", null.ok = TRUE)
  pdata
}

is_missing_prediction_data.PredictionDataVAE = function(pdata, ...) {
  pdata$row_ids[apply(pdata$response, 1L, anyNA)]
}

as_prediction.PredictionDataVAE = function(x, check = TRUE, ...) {
  invoke(PredictionVAE$new, check = check, .args = x)
}

c.PredictionDataVAE = function(..., keep_duplicates = TRUE) {
  dots = list(...)
  assert_list(dots, "PredictionDataVAE")
  if (length(dots) == 1L) return(dots[[1L]])

  row_ids = do.call(c, map(dots, "row_ids"))
  response = do.call(rbind, map(dots, "response"))
  if (!keep_duplicates) {
    keep = !duplicated(row_ids, fromLast = TRUE)
    row_ids = row_ids[keep]
    response = response[keep, , drop = FALSE]
  }

  pdata = list(row_ids = row_ids, response = response)
  class(pdata) = c("PredictionDataVAE", "PredictionData")
  pdata
}

filter_prediction_data.PredictionDataVAE = function(pdata, row_ids, ...) {
  keep = pdata$row_ids %in% row_ids
  pdata$row_ids = pdata$row_ids[keep]
  pdata$response = pdata$response[keep, , drop = FALSE]
  pdata
}

as.data.table.PredictionVAE = function(x, ...) {
  tab = data.table(row_ids = x$data$row_ids)
  cbind(tab, setnames(as.data.table(x$data$response), paste0("response.", colnames(x$data$response))))
}
```

### A Measure that Reads the Task {#sec-vae-measure}

Because the prediction carries no `truth`, our measure cannot compare `prediction$truth` to `prediction$response` the way the Hamming loss did.
The ground truth is the input itself, so the measure has to go back to the task.
Declaring the property `"requires_task"` makes `mlr3` pass the task to `$.score()`:

```{r}
MeasureVAEReconstruction = R6Class("MeasureVAEReconstruction",
  inherit = Measure,
  public = list(
    initialize = function() {
      super$initialize(
        id = "vae.reconstruction",
        task_type = "vae",
        predict_type = "response",
        properties = "requires_task",
        range = c(0, Inf),
        minimize = TRUE,
        label = "Mean Squared Reconstruction Error"
      )
    }
  ),
  private = list(
    .score = function(prediction, task, ...) {
      truth = as.matrix(task$data(rows = prediction$row_ids, cols = task$feature_names))
      mean((truth - prediction$response)^2)
    }
  )
)

mlr_measures$add("vae.reconstruction", MeasureVAEReconstruction)
```

Note that the measure subsets the task by `prediction$row_ids` rather than assuming that the prediction covers the whole task.
This matters during resampling, where each iteration predicts on a different subset.

### The Network {#sec-vae-network}

The `mlr3torch` side needs the same three generics as before.
`output_dim_for()` is the number of features, because the network reconstructs its input:

```{r}
output_dim_for.TaskVAE = function(x, ...) {
  length(x$feature_names)
}
```

`encode_prediction()` turns the network output into a matrix with one column per feature:

```{r}
encode_prediction.TaskVAE = function(task, predict_tensor, predict_type, ...) {
  reconstruction = as.matrix(predict_tensor$cpu())
  colnames(reconstruction) = task$feature_names
  list(response = reconstruction)
}
```

We do *not* implement `get_target_batchgetter()` here.
That generic is only used by the default dataset construction, which reads the target from the task's target columns, and a `TaskVAE` has none.
We will construct the dataset ourselves in the section on the dataset and the learner instead.

Now for the network itself.
The encoder maps the input to the mean and the log-variance of a Gaussian in latent space, and the decoder maps a latent code back to the input space.
During training we sample a latent code with the reparameterization trick and return all three tensors, because the loss needs the mean and the log-variance for the KL term:

```{r}
nn_vae = nn_module("nn_vae",
  initialize = function(d_in, d_hidden, d_latent) {
    self$encoder = nn_sequential(nn_linear(d_in, d_hidden), nn_relu())
    self$mean = nn_linear(d_hidden, d_latent)
    self$log_var = nn_linear(d_hidden, d_latent)
    self$decoder = nn_sequential(
      nn_linear(d_latent, d_hidden), nn_relu(), nn_linear(d_hidden, d_in)
    )
  },
  forward = function(input) {
    h = self$encoder(input)
    mean = self$mean(h)
    log_var = self$log_var(h)

    if (!self$training) {
      # when predicting, reconstruct deterministically from the latent mean
      return(self$decoder(mean))
    }

    # reparameterization trick: sample z ~ N(mean, exp(log_var))
    z = mean + torch_exp(0.5 * log_var) * torch_randn_like(mean)
    list(self$decoder(z), mean, log_var)
  }
)
```

The two branches of `forward()` implement a contract that `LearnerTorch` documents in its *Network Head and Target Encoding* section.
A network may return a `list()` of tensors during training, of which the first element is the primary prediction, i.e. the one that `measures_train` scores.
When the network is in evaluation mode it has to return that primary prediction as a single tensor, because `mlr3torch` concatenates the per-batch outputs along the batch dimension.
For a VAE this distinction is natural rather than a workaround: sampling belongs to training, and a deterministic reconstruction from the latent mean is what we want at prediction time.

### The Loss {#sec-vae-loss}

The loss receives the complete network output, so it can pick the three tensors apart itself.
The negative ELBO is the sum of a reconstruction term and a KL term that pulls the latent distribution towards a standard normal, and `beta` weights the two against each other:

```{r}
nn_elbo_loss = nn_module("nn_elbo_loss",
  initialize = function(beta = 1) {
    self$beta = assert_number(beta, lower = 0)
  },
  forward = function(input, target) {
    reconstruction = nnf_mse_loss(input[[1L]], target, reduction = "sum")
    kl = -0.5 * torch_sum(1 + input[[3L]] - input[[2L]]^2 - torch_exp(input[[3L]]))
    (reconstruction + self$beta * kl) / target$shape[1L]
  }
)

mlr3torch_losses$add("elbo", TorchLoss$new(
  torch_loss = nn_elbo_loss,
  task_types = "vae",
  id = "elbo",
  label = "Negative ELBO"
))

t_loss("elbo")
```

Because `beta` is an argument of `nn_elbo_loss()`, `TorchLoss` picks it up automatically and exposes it to the user as the hyperparameter `loss.beta`.

Note that no wrapping of the loss is necessary here.
`LearnerTorch` offers a private `$.loss_fn()` method for networks whose output has to be reshaped before the configured loss can be applied, but our loss consumes the list of three tensors directly.

### The Dataset and the Learner {#sec-vae-learner}

A `LearnerTorch` builds its dataset either from `$.ingress_tokens()`, which is what the multi-label learner did, or from `$.dataset()`, which gives full control.
We need the latter, because the target of a batch is the batch's own input and no target column exists.

Rather than building the dataset from scratch, we reuse `task_dataset()` and add the missing `y` element.
This keeps the row-id bookkeeping and the `lazy_tensor` handling of `task_dataset()` intact:

```{r}
vae_dataset = dataset("vae_dataset",
  initialize = function(dataset) {
    self$dataset = dataset
  },
  .getbatch = function(index) {
    batch = self$dataset$.getbatch(index)
    # a VAE reconstructs its own input, so the target of a batch is its input
    batch$y = batch$x$input
    batch
  },
  .length = function() length(self$dataset)
)
```

The learner then plugs the network, the dataset and the default loss together:

```{r}
LearnerVAE = R6Class("LearnerVAE",
  inherit = LearnerTorch,
  public = list(
    initialize = function(optimizer = NULL, loss = t_loss("elbo"), callbacks = list()) {
      param_set = ps(
        d_hidden = p_int(lower = 1L, tags = c("train", "required")),
        d_latent = p_int(lower = 1L, tags = c("train", "required"))
      )
      param_set$set_values(d_hidden = 16L, d_latent = 2L)

      super$initialize(
        id = "vae",
        task_type = "vae",
        label = "Variational Autoencoder",
        param_set = param_set,
        optimizer = optimizer,
        loss = loss,
        callbacks = callbacks,
        feature_types = c("numeric", "integer"),
        man = "mlr3torch::mlr_learners_vae"
      )
    }
  ),
  private = list(
    .network = function(task, param_vals) {
      nn_vae(length(task$feature_names), param_vals$d_hidden, param_vals$d_latent)
    },
    .dataset = function(task, param_vals) {
      ingress = list(input = ingress_num(shape = c(NA, length(task$feature_names))))
      vae_dataset(task_dataset(task, feature_ingress_tokens = ingress))
    }
  )
)
```

### Training and Predicting {#sec-vae-train}

From here on nothing is special anymore.
We train on two thirds of the data and reconstruct the held-out third:

```{r}
lrn_vae = LearnerVAE$new()
lrn_vae$param_set$set_values(epochs = 100, batch_size = 32, opt.lr = 0.01)

splits_iris = partition(tsk_iris_unsup)
lrn_vae$train(tsk_iris_unsup, row_ids = splits_iris$train)
prediction_vae = lrn_vae$predict(tsk_iris_unsup, row_ids = splits_iris$test)
prediction_vae
```

The prediction has no `truth.*` columns, because the task is unsupervised.
Scoring therefore needs the task, which we pass to `$score()`:

```{r}
prediction_vae$score(msr("vae.reconstruction"), task = tsk_iris_unsup)
```

The features were standardized to unit variance, so a mean squared reconstruction error well below `1` means that a two-dimensional latent space captures most of the structure of the four original features.

`resample()` passes the task on by itself, so no extra argument is needed there:

```{r}
rr_vae = resample(tsk_iris_unsup, lrn_vae, rsmp("cv", folds = 3))
rr_vae$aggregate(msr("vae.reconstruction"))
```

Validation and early stopping work as well, and are configured exactly as for multi-label classification:

```{r}
lrn_vae_es = LearnerVAE$new()
lrn_vae_es$param_set$set_values(
  epochs = 200, batch_size = 32, opt.lr = 0.01,
  patience = 10, measures_valid = msr("vae.reconstruction")
)
lrn_vae_es$validate = 0.3

lrn_vae_es$train(tsk_iris_unsup)
lrn_vae_es$internal_valid_scores
lrn_vae_es$internal_tuned_values
```

Finally, the trained network is available through `$network`, so the latent codes -- arguably the more interesting output of a VAE -- are only an encoder call away:

```{r}
network = lrn_vae$network
network$eval()

x = torch_tensor(
  as.matrix(tsk_iris_unsup$data(rows = splits_iris$test)),
  dtype = torch_float()
)
latent = with_no_grad(network$mean(network$encoder(x)))
head(as.matrix(latent))
```

If you want these codes to be a proper prediction rather than a manual encoder call, add `"latent"` to `mlr_reflections$learner_predict_types$vae`, give `PredictionVAE` a `$latent` binding, and return both matrices from `encode_prediction()`.
The one thing to keep in mind is that the network still has to return a *single* tensor when predicting, so the reconstruction and the latent code have to be concatenated in `forward()` and split apart again in `encode_prediction()`.

## Summary {#sec-summary}

Adding a task type to `mlr3torch` means writing two groups of things.

On the `mlr3` side:

* A row in `mlr_reflections$task_types`, plus the entries in `task_col_roles`, `task_properties`, `learner_properties`, `learner_predict_types`, `measure_properties` and `default_measures`.
* A `Task` class, inheriting from `TaskSupervised` or from `TaskUnsupervised`.
* A `Prediction` class and the S3 methods `check_prediction_data()`, `is_missing_prediction_data()`, `as_prediction()`, `c()`, `filter_prediction_data()` and `as.data.table()`.
* At least one `Measure`. If the task is unsupervised, give it the property `"requires_task"` so that it can read the ground truth from the task.

On the `mlr3torch` side:

* `output_dim_for()`, which says how wide the output layer of the network is.
* `get_target_batchgetter()`, which says how the target is turned into a tensor. A learner that builds its dataset through the private `$.dataset()` method does not need it.
* `encode_prediction()`, which says how the network output is turned back into a prediction.
* A `TorchLoss` that lists the new task type in its `task_types`.

Once these exist, `LearnerTorch` and everything built on it -- `resample()`, `benchmark()`, validation, early stopping, internal tuning and the `nn()` graph language -- works for the new task type without further changes.
