---
title: "Custom Learning Problems"
author: "Sebastian Fischer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Custom Learning Problems}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r custom_task_type-001, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lgr::get_logger("mlr3")$set_threshold("warn")
```

This article will show how to use {mlr3torch} for tasks that go beyond the classification and regression tasks that are supported out of the box.
This is possible via the generic `TaskTorch`, which allows you to use mlr3torch with arbitrary modeling problems.
The price of this flexibility is fewer compatibility checks and thus more responsibility on the user, see [Drawbacks of `TaskTorch`](#sec-drawbacks) for more information.
In this article we will first show how to use `TaskTorch` in a multi-label classification problem and then how it can be used for an unsupervised autoencoder.

## Multi-label Classification

As an example, we will train a learner on a multi-label classification problem which consists of $k$ binary prediction problems where multiple labels can be true.
Below, we generate some synthetic data for it and want to predict from the five `x` features whether it's sunny, warm, and windy.

```{r custom_task_type-002}
library(mlr3torch)
library(data.table)

set.seed(314)
n = 500
dat = data.table(x1 = rnorm(n), x2 = rnorm(n), x3 = rnorm(n), x4 = rnorm(n), x5 = rnorm(n))
dat[, `:=`(
  sunny = x1 + x2 > 0,
  warm = x2 * x3 > 0,
  windy = x4 - x5 > 0
)]
```

In order to create a `TaskTorch` for this modeling problem, we want to first specify:

1. The default method for converting torch predictions to an `mlr3::Prediction` object,
2. The default `mlr3::Measure` for scoring these predictions, and
3. How many output units a network for this task needs.

For our problem, we want to accept both response (class) predictions and probability predictions, and we assume that the network outputs per-class logits.
Any learner that we create for this task can also deviate from this default encoding.

```{r custom_task_type-003}
weather_encoder = function(task, network_output, predict_type) {
  prob = as.matrix(nnf_sigmoid(network_output)$cpu())
  colnames(prob) = task$target_names
  list(response = prob > 0.5, prob = if (predict_type == "prob") prob)
}
```

For more information on what a prediction may hold and how it is tabled, see `?PredictionTorch`.
As the default measure, we use the *Hamming loss* and construct it via `msr_torch`, see its help page for more information.

```{r custom_task_type-004}
msr_hamming = msr_torch("multilabel.hamming",
  function(truth, response) mean(as.matrix(truth) != response),
  range = c(0, 1), minimize = TRUE
)
msr_hamming
```

Below, we create the `TaskTorch`, where the third of these is the `output_dim` argument, a function of the task.

```{r custom_task_type-005}
tsk_weather = as_task_torch(dat, target = c("sunny", "warm", "windy"), id = "weather",
  output_dim = function(task) length(task$target_names),
  default_encoder = weather_encoder,
  default_measure = msr_hamming
)
tsk_weather

tsk_weather$truth(1:5)
tsk_weather$target_names
tsk_weather$feature_names
```

The `output_dim` field is used via the `output_dim_for()` generic:

```{r custom_task_type-006}
output_dim_for(tsk_weather)
```

Below, we define the architecture where the output dimension is obtained using `output_dim_for()`.

```{r custom_task_type-007}
nn_mlp = nn_module("nn_mlp",
  initialize = function(task, latent) {
    self$net = nn_sequential(
      nn_linear(length(task$feature_names), latent), nn_relu(),
      nn_linear(latent, latent), nn_relu(),
      nn_linear(latent, output_dim_for(task))
    )
  },
  # the argument name matches the name of the ingress token below
  forward = function(input) self$net(input)
)
```

To turn this into a learner, we additionally need a loss and a way to construct the batches.
For the features, we use a standard numeric encoder via `ingress_num()`, so we only specify the `target_batchgetter` to also encode the targets numerically.
The data argument of the batchgetter is `task$data(batch_ids, task$target_names)` and it must return a torch tensor.
For the loss, we use the binary cross entropy, built from the corresponding torch loss:

```{r custom_task_type-008}
loss_bce = as_torch_loss(nn_bce_with_logits_loss, id = "bce")
loss_bce

# the loss wants a float tensor of zeros and ones, so that is what the learner builds
weather_batchgetter = function(data) torch_tensor(as.matrix(data), dtype = torch_float())

lrn_weather = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  predict_types = c("response", "prob"),
  latent = 32,
  epochs = 5, batch_size = 32, opt.lr = 0.01
)
lrn_weather$predict_type = "prob"
```

Next, we train the learner on two thirds of the data and make predictions on the remaining observations:

```{r custom_task_type-009}
split = partition(tsk_weather)
lrn_weather$train(tsk_weather, split$train)
pred = lrn_weather$predict(tsk_weather, split$test)
pred
```

The prediction is scored like any other.
Note that a `mlr3::Measure` differs from a `TorchLoss`: it is computed from the encoded prediction, whereas the torch loss operates on the raw network output and the tensor-encoded target.

```{r custom_task_type-010}
pred$score(msr_hamming)
```

It is also possible to get the raw network outputs as a `lazy_tensor`:

```{r custom_task_type-011}
lrn_logits = lrn_weather$clone(deep = TRUE)
lrn_logits$predict_type = "lazy_tensor"

pred_logits = lrn_logits$predict(tsk_weather, row_ids = split$test[1:3])
pred_logits
materialize(pred_logits$lazy_tensor, rbind = TRUE)
```

Of course, we can also resample the learner.

```{r custom_task_type-012}
rr = resample(tsk_weather, lrn_weather, rsmp("cv", folds = 3))
rr$aggregate()
```

We can also construct the learner as a `Graph`.
Here, we have to specify the `target_batchgetter` in the `PipeOpTorchModel`.
Note that `nn("head")` also relies on `output_dim_for` to define the last layer.

```{r custom_task_type-013}
architecture = po("torch_ingress_num") %>>%
  nn("linear_1", out_features = 20) %>>%
  nn("relu_1") %>>%
  nn("linear_2", out_features = 20) %>>%
  nn("relu_2") %>>%
  nn("head") %>>%
  po("torch_loss", loss_bce) %>>%
  po("torch_optimizer", "adam", lr = 0.01) %>>%
  po("torch_model", batch_size = 32, epochs = 50,
    target_batchgetter = weather_batchgetter)

glrn_weather = as_learner(architecture)
glrn_weather$train(tsk_weather, row_ids = split$train)
glrn_weather$predict(tsk_weather, row_ids = split$test)$score(msr_hamming)
```

## A Simple Autoencoder {#sec-unsupervised}

A `TaskTorch` is supervised or unsupervised depending only on whether you gave it target columns.
Nothing beyond that is assumed about the structure of the problem.

A task with no targets has no target element in its batches at all, so the loss is called as `loss(y_hat)`, with no second argument.

If the target of a batch is a function of its *input*, as it is for an autoencoder reconstructing its input, a denoising or masked objective, or contrastive pretraining, then the learner's `target_batchgetter` may declare an `x` argument, which receives the feature tensors of the batch.
An autoencoder over the numeric features of `iris` is then:

```{r custom_task_type-014}
iris_scaled = as.data.table(scale(iris[, 1:4]))

tsk_ae = as_task_torch(iris_scaled, id = "iris_ae",
  output_dim = function(task) length(task$feature_names),
  default_encoder = function(task, network_output, predict_type) {
    response = as.matrix(network_output$cpu())
    colnames(response) = task$feature_names
    list(response = response)
  }
)
tsk_ae
```

Such a task has no `truth`, so its measure reads the ground truth from the task.
`msr_torch()` arranges that for any function that declares a `task` argument:

```{r custom_task_type-015}
msr_recon = msr_torch("reconstruction", function(task, prediction) {
  truth = as.matrix(task$data(rows = prediction$row_ids, cols = task$feature_names))
  mean((truth - prediction$response)^2)
}, range = c(0, Inf))
```

The network is an ordinary autoencoder, and the loss compares its output to the target that the batchgetter produced:

```{r custom_task_type-016}
nn_ae = nn_module("nn_ae",
  initialize = function(task) {
    d_in = length(task$feature_names)
    self$encoder = nn_sequential(
      nn_linear(d_in, 16),
      nn_relu(),
      nn_linear(16, 2)
    )
    self$decoder = nn_sequential(
      nn_linear(2, 16),
      nn_relu(),
      nn_linear(16, d_in)
    )
  },
  forward = function(input) self$decoder(self$encoder(input))
)

lrn_ae = lrn("torch.module",
  module_generator = nn_ae,
  ingress_tokens = list(input = ingress_num()),
  loss = t_loss("mse"),
  # a reconstruction objective: the target of a batch is the batch's own input
  target_batchgetter = function(data, x) x[[1L]],
  epochs = 100, batch_size = 32, opt.lr = 0.01
)

lrn_ae$train(tsk_ae)
lrn_ae$predict(tsk_ae)$score(msr_recon, task = tsk_ae)
```

### Predictions as Tensors {#sec-raw-predictions}

In some cases, you mighth want to directly access the raw tensor predictions instead of an encoded variant.
This is possible via the `"lazy_tensor"` predict type:

```{r custom_task_type-017}
lrn_raw = lrn_ae$clone(deep = TRUE)
lrn_raw$predict_type = "lazy_tensor"

pred_raw = lrn_raw$predict(tsk_ae)
pred_raw
```

You can convert this to an actual `torch_tensor` by materializing it:

```{r custom_task_type-018}
reconstruction = materialize(pred_raw$lazy_tensor, rbind = TRUE)
reconstruction$shape
```

Note that this prediction can currently not be correctly saved via `saveRDS()` as it holds external pointers.

You can also directly access the network, e.g. to get the latent embeddings:

```{r custom_task_type-020}
ds = lrn_ae$dataset(tsk_ae)
batch = ds$.getbatch(1:5)

network = lrn_ae$network
network$eval()
with_no_grad(network$encoder(batch$x$input))$cpu()
```

## Drawbacks of `TaskTorch` {#sec-drawbacks}

Now that we have shown the capabilities of `TaskTorch`, it is also important to highlight the drawbacks of the approach.
In {mlr3}, objects are annotated with meta-information that describes their properties and capabilities.
For example, training a classification learner on a regression task raises a clear error message:

```{r custom_task_type-021, error = TRUE}
lrn("classif.featureless")$train(tsk("mtcars"))
```

This is because they have different `task_type`s.
Because the `"torch"` task type can represent an arbitrary task type (that is what it is designed for), we can't distinguish between the multi-label classification task and the reconstruction problem.
E.g., training the autoencoder on the weather data is simply undefined behavior.
The same also holds true for measures.
As such, the `TaskTorch` is a *quick and dirty* way to approach a modeling problem.
The alternative is to properly create a new task type, which requires implementing sub-classes (such as `TaskRegr`, `MeasureRegr`, etc.) and registering the new task type with {mlr3}.
