---
title: "Custom Learning Problems"
author: "Sebastian Fischer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Custom Learning Problems}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r custom_task_type-001, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lgr::get_logger("mlr3")$set_threshold("warn")
```

This article will show how to use {mlr3torch} for tasks that go beyond the classification and regression tasks that are supported out of the box.
This is possible via the generic `TaskTorch`, which allows you to use mlr3torch with arbitrary modeling problems.
The price of this flexibility is fewer compatibility checks and thus more responsibility on the user, see @sec-drawbacks for more information.
In this article we will first show how to use `TaskTorch` in a multi-label classification problem and then how it can be used for an unsupervised autoencoder.

## Multi-label Classification

As an example, we will train a learner on a multi-label classification problem which consists of $k$ binary prediction problems where multiple labels can be true.
Below, we generate some synthetic data for it.

```{r custom_task_type-002}
library(mlr3torch)
library(data.table)

set.seed(314)
n = 500
dat = data.table(x1 = rnorm(n), x2 = rnorm(n), x3 = rnorm(n), x4 = rnorm(n), x5 = rnorm(n))
dat[, `:=`(
  sunny = x1 + x2 > 0,
  warm = x2 * x3 > 0,
  windy = x4 - x5 > 0
)]
```

Next, we create a `TaskTorch` from the data using `as_task_torch()`.
We also provide a default prediction encoder that describes how the raw torch predictions of learners are converted into a prediction object.
For our problem, we want to accept both response (class) predictions and probability predictions, and we assume that the network outputs per-class logits.
Any learner that we create for this task can also deviate from this default encoding.

```{r custom_task_type-003}
weather_encoder = function(task, network_output, predict_type) {
  prob = as.matrix(nnf_sigmoid(network_output)$cpu())
  colnames(prob) = task$target_names
  list(response = prob > 0.5, prob = if (predict_type == "prob") prob)
}
```

For more information on what a prediction may hold and how it is tabled, see `?PredictionTorch`.

The task also carries the measure that `mlr3` falls back to when none is named, because `mlr3torch` cannot know what a sensible default is for a problem it knows nothing about.
Ours is the *Hamming loss*, the fraction of individual label decisions that are wrong, written as a plain R function and wrapped with `msr_torch()`:

```{r custom_task_type-004}
msr_hamming = msr_torch("multilabel.hamming",
  function(truth, response) mean(as.matrix(truth) != response),
  range = c(0, 1), minimize = TRUE
)
msr_hamming
```

Below, we also specify the output dimension that the head of a network for this task has.
Note that the output dimension describes a single tensor: a network with more than one head sizes them itself, and a task for such a network simply does not set `output_dim`.

```{r custom_task_type-005}
tsk_weather = as_task_torch(dat, target = c("sunny", "warm", "windy"), id = "weather",
  output_dim = function(task) length(task$target_names),
  default_encoder = weather_encoder,
  default_measure = msr_hamming
)
tsk_weather

tsk_weather$truth(1:5)
tsk_weather$target_names
tsk_weather$feature_names
```

The output dimension for a task can be retrieved as follows:

```{r custom_task_type-006}
output_dim_for(tsk_weather)
```

Next, we will create a learner for the task.
Below, we define the architecture where the output dimension is obtained using `output_dim_for()`.

```{r custom_task_type-007}
nn_mlp = nn_module("nn_mlp",
  initialize = function(task, latent) {
    self$net = nn_sequential(
      nn_linear(length(task$feature_names), latent), nn_relu(),
      nn_linear(latent, latent), nn_relu(),
      nn_linear(latent, output_dim_for(task))
    )
  },
  # the argument name matches the name of the ingress token below
  forward = function(input) self$net(input)
)
```

To turn this into a learner, we additionally need a loss and a way to construct the batches.
For the features, we use a standard numeric encoder via `ingress_num()`, so we only specify the `target_batchgetter` to also encode the targets numerically.
The data argument of the batchgetter is `task$data(batch_ids, task$target_names)` and it must return a torch tensor.
For the loss, we use the binary cross entropy, built from the corresponding torch loss:

```{r custom_task_type-008}
loss_bce = as_torch_loss(nn_bce_with_logits_loss, id = "bce")

# the loss wants a float tensor of zeros and ones, so that is what the learner builds
weather_batchgetter = function(data) torch_tensor(1 * as.matrix(data), dtype = torch_float())

lrn_weather = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  predict_types = c("response", "prob", "lazy_tensor"),
  # `latent` is an argument of `nn_mlp()`, which makes it a hyperparameter of the learner
  latent = 32,
  epochs = 5, batch_size = 32, opt.lr = 0.01
)
lrn_weather$predict_type = "prob"
```

Next, we train the learner on two thirds of the data and make predictions on the remaining observations:

```{r custom_task_type-009}
split = partition(tsk_weather)
lrn_weather$train(tsk_weather, split$train)
pred = lrn_weather$predict(tsk_weather, split$test)
pred
```

The prediction is scored like any other.
Note that a `mlr3::Measure` differs from a `TorchLoss`: it is computed from the encoded prediction, whereas the torch loss operates on the raw network output and the tensor-encoded target.

```{r custom_task_type-010}
pred$score(msr_hamming)
```

Of course, we can also resample the learner. Since the task carries a default measure, `$aggregate()` needs no argument:

```{r custom_task_type-011}
rr = resample(tsk_weather, lrn_weather, rsmp("cv", folds = 3))
rr$aggregate()
```

## Scoring {#sec-measure}

A measure for a `TaskTorch` is a plain R function wrapped by `msr_torch()`.
The function receives whichever of the arguments `truth`, `response`, `prob`, `se`, `lazy_tensor`, `prediction`, `task`, `learner`, `train_set` and `weights` it declares, and returns a single number.
Asking for anything else is an error, and an argument that this prediction does not have -- `prob` when only a response was predicted, `weights` on a task without a `weights_measure` column -- is not passed at all, so a default you declare for it is what your function sees.

What `truth` and `response` look like is decided by the task and its encoder, not by this class.
For our task `truth` is a `data.table` with one logical column per label and `response` is a logical matrix, which is why `msr_hamming` above compares them with `as.matrix(truth) != response`.

A measure that consumes probabilities instead only has to say so with `predict_type = "prob"` and declare a `prob` argument.
The mean binary log-loss over all labels is one example:

```{r custom_task_type-012}
msr_logloss = msr_torch("multilabel.logloss",
  function(truth, prob) {
    eps = 1e-7
    truth = as.matrix(truth)
    prob = pmin(pmax(prob, eps), 1 - eps)
    -mean(truth * log(prob) + (1 - truth) * log(1 - prob))
  },
  predict_type = "prob", range = c(0, Inf), minimize = TRUE
)
```

Declaring a `learner` argument gets you the learner object, which for a torch measure usually means the network.
Note that this only adds the `"requires_learner"` property, and `mlr3` also passes a learner whose model was thrown away -- after a `resample()` with `store_models = FALSE`, `learner$network` is `NULL` and the measure quietly scores an empty model.
A measure that reaches for the network should therefore ask for the model too, with `properties = "requires_model"`, so that scoring without one is an error.

## Training a Network {#sec-train}

Networks are supplied to `lrn("torch.module")`, the counterpart of `lrn("classif.module")` and `lrn("regr.module")` for the new task type.
The module generator receives the `task`, so it can size its layers from the task itself -- in particular its output layer, through `output_dim_for()`.
The `ingress_tokens` argument says how the *features* enter the network; `ingress_num()` collects all numeric features into one tensor.
Since the network emits one logit per label, the matching loss is the binary cross entropy, which we have to pass explicitly.

From here the learner behaves like any other `mlr3` learner:

```{r custom_task_type-013}
splits = partition(tsk_weather)
lrn_weather$train(tsk_weather, row_ids = splits$train)
prediction = lrn_weather$predict(tsk_weather, row_ids = splits$test)
prediction
```

The `truth.*` columns are the ground truth taken from the task, `response` holds what the network predicted, and the `prob.*` columns hold the probabilities we asked for.
Both of our measures work on it:

```{r custom_task_type-014}
prediction$score(c(msr_hamming, msr_logloss))
```

The encoder can also be skipped entirely.
The predict type `"lazy_tensor"` hands back what the network produced -- here the three logits per observation, before anything turned them into labels or probabilities -- which is why `lrn_weather` was constructed with it among its `predict_types`:

```{r custom_task_type-015}
lrn_logits = lrn_weather$clone(deep = TRUE)
lrn_logits$predict_type = "lazy_tensor"

pred_logits = lrn_logits$predict(tsk_weather, row_ids = splits$test)
pred_logits
materialize(pred_logits$lazy_tensor, rbind = TRUE)[1:3, ]
```

@sec-raw-predictions comes back to this predict type and to what you give up by using it.

Resampling works too.
This is the point at which a hand-written task type would already have needed its `PredictionData` methods, because the predictions of the individual folds have to be combined and subset again:

```{r custom_task_type-016}
rr = resample(tsk_weather, lrn_weather, rsmp("cv", folds = 3))
rr$aggregate(msr_hamming)
```

The network does not have to be a module generator: everything in the [Defining an Architecture](pipeop_torch.html) article applies here as well, with `po("torch_model")` as the terminal `PipeOp` for a `TaskTorch`, alongside the `po("torch_model_classif")` and `po("torch_model_regr")` you may already know.

```{r custom_task_type-017}
architecture = po("torch_ingress_num") %>>%
  nn("linear_1", out_features = 20) %>>%
  nn("relu_1") %>>%
  nn("linear_2", out_features = 20) %>>%
  nn("relu_2") %>>%
  nn("head") %>>%
  po("torch_loss", loss_bce) %>>%
  po("torch_optimizer", "adam", lr = 0.01) %>>%
  po("torch_model", batch_size = 32, epochs = 50,
    target_batchgetter = weather_batchgetter)

glrn_weather = as_learner(architecture)
glrn_weather$train(tsk_weather, row_ids = splits$train)
glrn_weather$predict(tsk_weather, row_ids = splits$test)$score(msr_hamming)
```

`nn("head")` sized its output layer through `output_dim_for()` and the resulting learner decoded its predictions through the task's prediction encoder.
Neither of them knows anything about multi-label classification.

### Validation and Early Stopping {#sec-validation}

Custom measures can be used for validation and early stopping exactly like built-in ones.
Below we hold out 30% of the training data, score it after every epoch, and stop once the Hamming loss has not improved for five epochs:

```{r custom_task_type-018}
lrn_early = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  latent = 32,
  epochs = 100, batch_size = 32, opt.lr = 0.01,
  patience = 5, measures_valid = msr_hamming
)
lrn_early$validate = 0.3

lrn_early$train(tsk_weather)
lrn_early$internal_valid_scores
lrn_early$internal_tuned_values
```

Training stopped well before the configured 100 epochs, and `$internal_tuned_values` reports the epoch that `mlr3tuning` would use for internal tuning.

## A Simple Autoencoder {#sec-unsupervised}

A `TaskTorch` is supervised or unsupervised depending only on whether you gave it target columns.
Nothing beyond that is assumed about the structure of the problem.

A task with no targets has no target element in its batches at all, so the loss is called as `loss(y_hat)`, with no second argument.

If the target of a batch is a function of its *input* -- an autoencoder reconstructing its input, a denoising or masked objective, contrastive pretraining -- then the learner's `target_batchgetter` may declare an `x` argument, which receives the feature tensors of the batch.
An autoencoder over the numeric features of `iris` is then:

```{r custom_task_type-019}
iris_scaled = as.data.table(scale(iris[, 1:4]))

tsk_ae = as_task_torch(iris_scaled, id = "iris_ae",
  output_dim = function(task) length(task$feature_names),
  default_encoder = function(task, network_output, predict_type) {
    response = as.matrix(network_output$cpu())
    colnames(response) = task$feature_names
    list(response = response)
  }
)
tsk_ae
```

Such a task has no `truth`, so its measure reads the ground truth from the task.
`msr_torch()` arranges that for any function that declares a `task` argument:

```{r custom_task_type-020}
msr_recon = msr_torch("reconstruction", function(task, prediction) {
  truth = as.matrix(task$data(rows = prediction$row_ids, cols = task$feature_names))
  mean((truth - prediction$response)^2)
}, range = c(0, Inf))
```

The network is an ordinary autoencoder, and the loss compares its output to the target that the batchgetter produced:

```{r custom_task_type-021}
nn_ae = nn_module("nn_ae",
  initialize = function(task) {
    d_in = length(task$feature_names)
    self$encoder = nn_sequential(nn_linear(d_in, 16), nn_relu(), nn_linear(16, 2))
    self$decoder = nn_sequential(nn_linear(2, 16), nn_relu(), nn_linear(16, d_in))
  },
  forward = function(input) self$decoder(self$encoder(input))
)

lrn_ae = lrn("torch.module",
  module_generator = nn_ae,
  ingress_tokens = list(input = ingress_num()),
  loss = t_loss("mse"),
  # a reconstruction objective: the target of a batch is the batch's own input
  target_batchgetter = function(data, x) x[[1L]],
  # `"prob"`, `"se"` and `"lazy_tensor"` are opt-in, see @sec-raw-predictions
  predict_types = c("response", "lazy_tensor"),
  epochs = 100, batch_size = 32, opt.lr = 0.01
)

lrn_ae$train(tsk_ae)
lrn_ae$predict(tsk_ae)$score(msr_recon, task = tsk_ae)
```

The features were standardized to unit variance, so a mean squared reconstruction error well below `1` means that a two-dimensional latent space captures most of the structure of the four original features.

Resampling works here too, and the task is passed on automatically:

```{r custom_task_type-022}
resample(tsk_ae, lrn_ae, rsmp("cv", folds = 3))$aggregate(msr_recon)
```

### Predictions as Tensors {#sec-raw-predictions}

The encoder above exists to turn the network's output into something a measure can be written against.
When the output *is* what you want, the predict type `"lazy_tensor"` hands it over directly, and the task then needs no `default_encoder` at all:

```{r custom_task_type-023}
lrn_raw = lrn_ae$clone(deep = TRUE)
lrn_raw$predict_type = "lazy_tensor"

pred_raw = lrn_raw$predict(tsk_ae)
pred_raw
```

One element per observation, holding that observation's reconstruction.
This predict type needs a network that returns a single tensor -- one element per observation is one tensor per observation, so there is no room for a second head.
`materialize()` turns them back into a tensor:

```{r custom_task_type-024}
reconstruction = materialize(pred_raw$lazy_tensor, rbind = TRUE)
reconstruction$shape
```

It is the same network output the encoder saw, just not encoded into anything:

```{r custom_task_type-025}
encoded = lrn_ae$predict(tsk_ae)$response
all.equal(as.matrix(reconstruction$cpu()), unname(encoded), tolerance = 1e-6)
```

This is the one predict type whose predictions do **not** survive `saveRDS()`.
They hold `torch` tensors, which are external pointers: saving succeeds and the tensors are invalid the next time they are touched, in this session or another -- a `ResampleResult` containing them keeps its row ids and scores but loses its tensors.
Keep the encoder for anything you want to store, and reach for `"lazy_tensor"` when you are working with the output there and then.

The latent codes -- arguably the more interesting output of an autoencoder -- are not a prediction at all, and are available through the trained network:

```{r custom_task_type-026}
network = lrn_ae$network
network$eval()
x = torch_tensor(as.matrix(tsk_ae$data(rows = 1:5)), dtype = torch_float())
as.matrix(with_no_grad(network$encoder(x)))
```

## Drawbacks of `TaskTorch` {#sec-drawbacks}


All of this convenience comes from one decision: every `TaskTorch` has the same task type, `"torch"`.
`mlr3` therefore cannot tell one custom problem from another.

```{r custom_task_type-027}
tsk_weather$task_type
tsk_ae$task_type
```

The consequences are worth being explicit about:

* A learner built for one problem will accept a task from a completely different one. If the tensor shapes happen to line up, it trains, and you get a meaningless model rather than an error.
* The same holds for measures: `mlr3` will not stop you scoring a multi-label prediction with a reconstruction error.
* There is no dictionary entry for your problem. `lrn("torch.module")` and `po("torch_model")` are shared by everyone, so you cannot hand a colleague a learner id and have it mean your problem.
* Encapsulation has nothing to fall back to. `mlr3` requires a fallback learner for `encapsulate = "evaluate"` or `"callr"` and looks one up by task type; nothing is registered for `"torch"`, so `resample(task, learner, resampling, encapsulate = "evaluate")` errors with *Could not find default fallback learner* instead of carrying on past a failing fold. Pass a fallback of your own to `learner$encapsulate()` if you need one.
* Not every `PipeOp` accepts a `TaskTorch`, see below.

## Summary

Before `mlr3torch` can train a network on a task it has to know three things, and a `TaskTorch` infers **none** of them:

| what | who provides it | how |
| --- | --- | --- |
| the target tensor `y` of a batch | the **learner** | `target_batchgetter` |
| the number of output units | the **task** | `output_dim`, read through `output_dim_for()` |
| how a network output becomes a prediction | the **task** | `default_encoder`, read through `encode_prediction()` |

Nothing is guessed because the column types of a target say how it is *stored*, not how you want to *model* it: a two-level `factor` could be one logit or two, and `d` numeric columns could be one head of width `d` or `d` separate heads.

The split between task and learner follows what each piece is consumed by.
What `y` has to look like is dictated by the loss, which belongs to the learner -- cross entropy wants `long` class indices where mean squared error wants `float`, so the same task trained with two losses needs two different target tensors.
What a prediction looks like is what the task's measures are written against, so it belongs to the task.

`output_dim` is a `function(task)` rather than a number so that it still holds after the task changes -- a different target column, a `$droplevels()`.
It is also optional: only `nn("head")` and modules that call `output_dim_for()` themselves ask for it, so a network that sizes its own output never needs it.

`default_encoder` is a *default* in the same sense as `default_measure`: it is what a learner uses when it has none of its own.
It receives the network's output unchanged, which makes a network with more than one head its business too, and it is where a consistency check between network and task belongs -- a network trained on three target levels emits three columns, and mapping those onto the levels of a task that has since been `$droplevels()`ed would relabel every observation rather than fail.
A `LearnerTorch` subclass may overwrite the private `$.encode_prediction()` method and ignore the task's encoder entirely; `lrn("classif.tabm")` averages over its ensemble dimension that way, and the `torchvision` learners strip the auxiliary head of an Inception network first.
The predict type `"lazy_tensor"` (@sec-raw-predictions) skips the encoding altogether.

The loss and the measure are not inferred either, for the same reason: which ones are right depends on what your problem means, so you state them yourself.
