---
title: "Writing your own PipeOpTorch"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing your own PipeOpTorch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(mlr3torch)
```

`mlr3torch` ships a `PipeOp` for most of the layers that `torch` provides, but sooner or later you
will want a layer of your own in a `Graph`.
This article shows how to get there, from the shortest path to the full R6 class.

If you have not built a network from `PipeOp`s before, read
[Defining an Architecture](pipeop_torch.html) first; for what a `ModelDescriptor` is and how a
`Graph` of `PipeOp`s becomes a `nn_module`, see [Networks as Graphs](internals_pipeop_torch.html).

## What a `PipeOpTorch` has to know

A network in `mlr3torch` is assembled without any data flowing through it.
Instead of tensors, the `Graph` passes around a `ModelDescriptor` that carries the *shape* of the
tensors that will later be there.
That is what makes `po("nn_linear", out_features = 10)` possible: you say how wide the layer should
be, and `in_features` follows from whatever is connected to its input.

A `PipeOpTorch` therefore answers two questions that the module itself does not:

1. **What does the module need to know about its input?** `in_features` of `nn_linear()` is not
   something the user sets, it is read off the incoming shape.
2. **What shape does the module produce?** The next layer needs it, for the same reason.

Everything else -- building the module, adding it to the network, connecting the edges -- is
inherited.

## The short way: `pipeop_torch()`

Take a layer that scales each feature by a learned factor.
Written with `pipeop_torch()`, it looks exactly like `nn_module()` would, with two additions:
`initialize()` takes `shapes_in` -- question 1 -- and `shapes_out` answers question 2.

```{r}
PipeOpTorchScale = pipeop_torch("nn_scale",
  initialize = function(shapes_in, init = 1) {
    self$weight = nn_parameter(torch_full(tail(shapes_in[[1L]], 1L), init))
  },
  forward = function(input) input * self$weight,
  shapes_out = function(shapes_in) shapes_in
)
```

The remaining arguments of `initialize()` become the hyperparameters of the operator: `init` is
something the user sets, the number of features is not.

```{r}
po_scale = PipeOpTorchScale$new()
po_scale$param_set$ids()
po_scale$shapes_out(list(c(NA, 4)))
```

Note the `NA`: the batch dimension is unknown while the network is being built, and any other
dimension may be too -- the sequence length of a transformer, or the size of an image.
Both `initialize()` and `shapes_out()` see those `NA`s and have to cope with them, either by
carrying them through, as here, or by saying which ones they need.
The helpers `assert_known_dims()`, `assert_ndim()` and friends state which dimensions you actually
need, and report the offending `PipeOp` and shape when one of them is unknown.
Be permissive with the rest: rejecting a shape that would have worked is worse than a runtime error
later.

That is a usable operator:

```{r}
graph = po("torch_ingress_num") %>>% po_scale %>>% po("nn_head")
md = graph$train(tsk("iris"))[[1L]]
network = model_descriptor_to_module(md)
network
```

The module was built with four features, taken from the task:

```{r}
as.numeric(md$graph$pipeops$nn_scale$module$weight)
```

Besides `shapes_in`, `initialize()` can take the `task`, and `shapes_out()` can take the
`param_vals` and the `task` as well -- each is passed only if the function declares it.
That is how a layer whose width is a hyperparameter states its output shape:

```{r}
PipeOpTorchLinear2 = pipeop_torch("nn_linear2",
  initialize = function(shapes_in, out_features, bias = TRUE) {
    self$linear = nn_linear(tail(shapes_in[[1L]], 1L), out_features, bias)
  },
  forward = function(input) self$linear(input),
  shapes_out = function(shapes_in, param_vals) {
    list(c(head(shapes_in[[1L]], -1L), param_vals$out_features))
  }
)

PipeOpTorchLinear2$new(param_vals = list(out_features = 10))$shapes_out(list(c(NA, 4)))
```

`out_features` has no default, so it is a *required* hyperparameter and the operator refuses to be
used without it.
This is also how you wrap a module that already exists: call it from `initialize()`.

## Converting a module with `as_pipeop()`

A module can also carry the answers itself, in which case `as_pipeop()` converts it without any
further configuration.
The rules are the same -- `initialize()` may take `shapes_in` and `task`, `shapes_out()` states the
output shapes -- except that `shapes_out()` is written as a method of the module:

```{r}
nn_scale = nn_module("nn_scale",
  initialize = function(shapes_in, init = 1) {
    self$weight = nn_parameter(torch_full(tail(shapes_in[[1L]], 1L), init))
  },
  forward = function(input) input * self$weight,
  shapes_out = function(shapes_in) shapes_in
)

as_pipeop(nn_scale)$shapes_out(list(c(NA, 4)))
```

It is called before any instance of the module exists, so it cannot use `self`.
A module written for `torch` rather than for `mlr3torch` has no such method, and `as_pipeop()` says
so rather than guessing.

## Several channels

A module whose `forward()` takes more than one tensor becomes a `PipeOp` with more than one input
channel -- the channels are named after the arguments, and `shapes_in` is named after the channels.
Returning a list makes several output channels, which have to be declared with `out_channels`:

```{r}
PipeOpTorchSplit = pipeop_torch("nn_split",
  initialize = function(shapes_in, d_out1, d_out2) {
    self$linear1 = nn_linear(tail(shapes_in$input, 1L), d_out1)
    self$linear2 = nn_linear(tail(shapes_in$input, 1L), d_out2)
  },
  forward = function(input) {
    list(self$linear1(input), self$linear2(input))
  },
  out_channels = 2,
  shapes_out = function(shapes_in, param_vals) {
    list(c(head(shapes_in$input, -1L), param_vals$d_out1),
      c(head(shapes_in$input, -1L), param_vals$d_out2))
  }
)

po_split = PipeOpTorchSplit$new(param_vals = list(d_out1 = 10, d_out2 = 20))
po_split$shapes_out(list(c(NA, 4)))
```

`in_channels` and `out_channels` also take a count instead of names -- `in_channels = 0` for a
module whose `forward()` takes `...`, i.e. a *vararg* channel.

## The long way: inheriting from `PipeOpTorch`

`pipeop_torch()` writes an R6 class for you; when you want to ship an operator in a package, or
need control over its construction, write that class yourself.
The module is then an ordinary `torch` module that knows nothing about shapes,

```{r}
nn_scale = nn_module("nn_scale",
  initialize = function(n_features, init = 1) {
    self$weight = nn_parameter(torch_full(n_features, init))
  },
  forward = function(input) input * self$weight
)
```

and the two questions are the private methods `.shape_dependent_params()` and `.shapes_out()`:

```{r}
PipeOpTorchScale = R6::R6Class("PipeOpTorchScale",
  inherit = PipeOpTorch,
  public = list(
    initialize = function(id = "nn_scale", param_vals = list()) {
      param_set = paradox::ps(
        init = paradox::p_dbl(default = 1, tags = "train")
      )
      super$initialize(
        id = id,
        param_vals = param_vals,
        param_set = param_set,
        module_generator = nn_scale
      )
    }
  ),
  private = list(
    .shape_dependent_params = function(shapes_in, param_vals, task) {
      c(param_vals, list(n_features = tail(shapes_in[[1L]], 1L)))
    },
    .shapes_out = function(shapes_in, param_vals, task) shapes_in
  )
)

PipeOpTorchScale$new()$shapes_out(list(c(NA, 4)))
```

`.shape_dependent_params()` returns *all* arguments the module is built with, i.e. the user's
`param_vals` plus the inferred ones.
If the module cannot be built by calling the generator with a list of arguments at all, overload
`.make_module(shapes_in, param_vals, task)` instead and ignore both.

To make the operator available as `po("nn_scale")`, add it to the `PipeOp` dictionary, typically in
your package's `.onLoad()`:

```{r, eval = FALSE}
mlr3pipelines::mlr_pipeops$add("nn_scale", PipeOpTorchScale)
```

## Registering the operator with a learner

Nothing else is needed: a `Graph` that contains the new operator becomes a `Learner` the usual way.

```{r}
graph = po("torch_ingress_num") %>>%
  PipeOpTorchScale$new() %>>%
  po("nn_head") %>>%
  po("torch_loss", t_loss("cross_entropy")) %>>%
  po("torch_optimizer", t_opt("adam")) %>>%
  po("torch_model_classif", batch_size = 32, epochs = 1)

learner = as_learner(graph)
learner$id = "scaled_mlp"
learner
```

## Checking your work

Two things are worth testing, and both are cheap:

* `$shapes_out()` agrees with what the module actually produces, for shapes with and without `NA`s.
  `mlr3torch`'s own test suite sweeps every combination of unknown dimensions for every operator it
  ships; the helper that does it lives in `tests/testthat/helper_shape_inference.R`.
* the operator survives a round-trip through a `Graph`, i.e. `model_descriptor_to_module()` returns
  a network that runs on a tensor of the announced shape.

```{r}
po_scale = PipeOpTorchScale$new()
inferred = po_scale$shapes_out(list(c(NA, 4)))[[1L]]

network = model_descriptor_to_module(
  (po("torch_ingress_num") %>>% po_scale)$train(tsk("iris"))[[1L]]
)
observed = dim(with_no_grad(network(torch_ingress_num.input = torch_randn(2, 4))))

inferred
observed
```
