---
title: "Custom Learning Problems"
author: "Sebastian Fischer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Custom Learning Problems}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r custom_task_type-001, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lgr::get_logger("mlr3")$set_threshold("warn")
set.seed(314)
```

In this article, we will train neural networks on learning problems that are neither classification nor regression, using the general-purpose task type `TaskTorch`.

`mlr3torch` supports classification (`"classif"`) and regression (`"regr"`) out of the box, and between them they cover a large share of supervised learning.
Not every problem fits into them, though.
Multi-label classification attaches several binary labels to each observation, multi-output regression predicts several numbers at once, and an autoencoder has no target column at all.

In `mlr3`, each of these would ordinarily be a separate *task type*: a task type is the contract that says what the target of a task looks like, what a prediction for it consists of, and how such a prediction is scored.
Adding one properly means writing a `Task` class, a `Prediction` class, half a dozen `PredictionData` methods, a `Measure`, and a batch of entries in `mlr_reflections` -- a few hundred lines before you train anything.

`TaskTorch` is the shortcut.
`mlr3torch` registers a single task type `"torch"` once, and every custom problem becomes an *instance* of `TaskTorch` rather than a new class.
What would otherwise be S3 methods dispatching on your task class are arguments you pass to the task and the learner.
The result is that the multi-label problem below takes about five lines to set up instead of a few hundred.

```{r custom_task_type-002, message = FALSE}
library(mlr3torch)
library(data.table)
```

## A First Custom Task {#sec-first}

Multi-label classification attaches several binary labels to each observation, and the labels are not mutually exclusive.
A weather observation can be sunny and warm and windy, or sunny and warm but not windy.
We represent such a problem as a table with one `logical` column per label.

The three labels below depend on the five features in different ways, and `warm` depends on an *interaction* of two features, which a linear model could not pick up:

```{r custom_task_type-003}
n = 500
dat = data.table(x1 = rnorm(n), x2 = rnorm(n), x3 = rnorm(n), x4 = rnorm(n), x5 = rnorm(n))
dat[, `:=`(
  sunny = x1 + x2 > 0,
  warm = x2 * x3 > 0,
  windy = x4 - x5 > 0
)]
```

`as_task_torch()` turns this into a task, in the same way that `as_task_classif()` and `as_task_regr()` do for the built-in types.
The only thing it needs beyond the data is which columns are the target:

```{r custom_task_type-004}
weather_encoder = function(task, predict_tensor, predict_type) {
  prob = as.matrix(nnf_sigmoid(predict_tensor)$cpu())
  colnames(prob) = task$target_names
  list(response = prob > 0.5, prob = if (predict_type == "prob") prob)
}

tsk_weather = as_task_torch(dat, target = c("sunny", "warm", "windy"), id = "weather",
  output_dim = function(task) length(task$target_names),
  prediction_encoder = weather_encoder
)
tsk_weather
```

That is the whole task.
It behaves like any other `mlr3` task -- it has features, targets, row and column roles, and it can be filtered, subset and resampled.

### What You Have to Specify {#sec-inference}

Before `mlr3torch` can train a network on a task it has to know three things: how the target of a batch becomes a tensor, how many output units the network needs, and how the raw output of the network becomes a prediction.
A `TaskTorch` infers **none** of them.

That is deliberate. The column types of a target say how it is *stored*, not how you want to *model* it: a two-level `factor` could be one logit or two, and `d` numeric columns could be one head of width `d` or `d` separate heads. Guessing would be a modelling decision made on your behalf, so you state all three yourself.

They do not all live in the same place:

| what | who provides it | how |
| --- | --- | --- |
| the target tensor `y` of a batch | the **learner** | `target_batchgetter` |
| the number of output units | the **task** | `output_dim`, read through `output_dim_for()` |
| how a network output becomes a prediction | the **task** | `prediction_encoder`, read through `encode_prediction()` |

The split follows what each one is consumed by.
What `y` has to look like is dictated by the loss, which belongs to the learner -- cross entropy wants `long` class indices where mean squared error wants `float`, so the same task trained with two losses needs two different target tensors.
What a prediction looks like is what the task's measures are written against, so it belongs to the task.

Our three `logical` columns are three independent binary decisions, so the network gets three output units:

```{r custom_task_type-005}
output_dim_for(tsk_weather)
```

`output_dim` is a `function(task)` rather than a number so that it still holds after the task changes -- a different target column, a `$droplevels()`.
It is optional: a network that sizes its own output, such as one with several heads of different widths, never asks for it, and only `nn("head")` and modules calling `output_dim_for()` do.

The prediction encoder receives the network's output unchanged, so a network with more than one head is its business too.
It is also where any consistency check between network and task belongs: a network trained on three target levels emits three columns, and mapping those onto the levels of a task that has since been `$droplevels()`ed would relabel every observation rather than fail.

The target tensor is the learner's, and we come to it in @sec-network.

> **Nothing is inferred, including the loss and the measure**
>
> There is no default loss and no default measure for a `TaskTorch` either: which ones are right depends on what your problem means, so you always state them yourself.

## Scoring {#sec-measure}

A measure for a `TaskTorch` is a plain R function wrapped by `msr_torch()`.
The function receives whichever of the arguments `truth`, `response`, `prob`, `prediction`, `task`, `learner` and `train_set` it declares, and returns a single number.

What `truth` and `response` look like is the last two columns of the table above.
For our task `truth` is a `data.table` with one logical column per label and `response` is a logical matrix, so the classic multi-label metric -- the *Hamming loss*, the fraction of individual label decisions that are wrong -- is a one-liner:

```{r custom_task_type-007}
msr_hamming = msr_torch("multilabel.hamming",
  function(truth, response) mean(as.matrix(truth) != response),
  range = c(0, 1), minimize = TRUE
)
msr_hamming
```

A measure that consumes probabilities instead only has to say so with `predict_type = "prob"` and declare a `prob` argument.
The mean binary log-loss over all labels is one example:

```{r custom_task_type-008}
msr_logloss = msr_torch("multilabel.logloss",
  function(truth, prob) {
    eps = 1e-7
    truth = as.matrix(truth)
    prob = pmin(pmax(prob, eps), 1 - eps)
    -mean(truth * log(prob) + (1 - truth) * log(1 - prob))
  },
  predict_type = "prob", range = c(0, Inf), minimize = TRUE
)
```

## Training a Network {#sec-train}

Networks are supplied to `lrn("torch.module")`, the counterpart of `lrn("classif.module")` and `lrn("regr.module")` for the new task type.
The module generator receives the `task`, so it can size its layers from the task itself -- in particular its output layer, through `output_dim_for()`:

```{r custom_task_type-009}
nn_mlp = nn_module("nn_mlp",
  initialize = function(task) {
    self$net = nn_sequential(
      nn_linear(length(task$feature_names), 20), nn_relu(),
      nn_linear(20, 20), nn_relu(),
      nn_linear(20, output_dim_for(task))
    )
  },
  # the argument name matches the name of the ingress token below
  forward = function(input) self$net(input)
)
```

The `ingress_tokens` argument says how the *features* enter the network; `ingress_num()` collects all numeric features into one tensor.
Since the network emits one logit per label, the matching loss is the binary cross entropy, which we have to pass explicitly:

```{r custom_task_type-010}
loss_bce = TorchLoss$new(nn_bce_with_logits_loss, id = "bce")

# the loss wants a float tensor of zeros and ones, so that is what the learner builds
weather_batchgetter = function(data) torch_tensor(1 * as.matrix(data), dtype = torch_float())

lrn_weather = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  epochs = 50, batch_size = 32, opt.lr = 0.01
)
lrn_weather$predict_type = "prob"
```

From here the learner behaves like any other `mlr3` learner:

```{r custom_task_type-011}
splits = partition(tsk_weather)
lrn_weather$train(tsk_weather, row_ids = splits$train)
prediction = lrn_weather$predict(tsk_weather, row_ids = splits$test)
prediction
```

The `truth.*` columns are the ground truth taken from the task and the `response.*` columns are what the network predicted.
Both of our measures work on it:

```{r custom_task_type-012}
prediction$score(c(msr_hamming, msr_logloss))
```

Resampling works too.
This is the point at which a hand-written task type would already have needed its `PredictionData` methods, because the predictions of the individual folds have to be combined and subset again:

```{r custom_task_type-013}
rr = resample(tsk_weather, lrn_weather, rsmp("cv", folds = 3))
rr$aggregate(msr_hamming)
```

### A Default Measure for the Task {#sec-default-measure}

`rr$aggregate()` without an argument falls back to the default measure of the task type, and `mlr3torch` cannot know what that should be for your problem.
You can attach one to the task, in which case `$score()` and `$aggregate()` work with no arguments at all:

```{r custom_task_type-014}
tsk_weather_m = as_task_torch(dat, target = c("sunny", "warm", "windy"), id = "weather",
  output_dim = function(task) length(task$target_names),
  prediction_encoder = weather_encoder,
  measure = msr_hamming)

resample(tsk_weather_m, lrn_weather, rsmp("holdout"))$aggregate()
```

Without it you get an error telling you to pass a measure explicitly.

### Validation and Early Stopping {#sec-validation}

Custom measures can be used for validation and early stopping exactly like built-in ones.
Below we hold out 30% of the training data, score it after every epoch, and stop once the Hamming loss has not improved for five epochs:

```{r custom_task_type-015}
lrn_early = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  epochs = 100, batch_size = 32, opt.lr = 0.01,
  patience = 5, measures_valid = msr_hamming
)
lrn_early$validate = 0.3

lrn_early$train(tsk_weather)
lrn_early$internal_valid_scores
lrn_early$internal_tuned_values
```

Training stopped well before the configured 100 epochs, and `$internal_tuned_values` reports the epoch that `mlr3tuning` would use for internal tuning.

## Building the Network as a Graph {#sec-graph}

Everything in the [Defining an Architecture](pipeop_torch.html) article applies here as well.
The terminal `PipeOp` for a `TaskTorch` is `po("torch_model")`, alongside the `po("torch_model_classif")` and `po("torch_model_regr")` you may already know:

```{r custom_task_type-016}
architecture = po("torch_ingress_num") %>>%
  nn("linear_1", out_features = 20) %>>%
  nn("relu_1") %>>%
  nn("linear_2", out_features = 20) %>>%
  nn("relu_2") %>>%
  nn("head") %>>%
  po("torch_loss", loss_bce) %>>%
  po("torch_optimizer", "adam", lr = 0.01) %>>%
  po("torch_model", batch_size = 32, epochs = 50,
    target_batchgetter = weather_batchgetter)

glrn_weather = as_learner(architecture)
glrn_weather$train(tsk_weather, row_ids = splits$train)
glrn_weather$predict(tsk_weather, row_ids = splits$test)$score(msr_hamming)
```

`nn("head")` sized its output layer through `output_dim_for()` and the resulting learner decoded its predictions through the task's prediction encoder.
Neither of them knows anything about multi-label classification.

## The Same Task, a Different Target Encoding {#sec-override}

Because the target tensor belongs to the learner, the same task can be trained two different ways without touching the task.
Suppose we want to treat a single `factor` target as a regression onto one-hot encoded labels rather than as multi-class classification.

```{r custom_task_type-017}
iris_dt = as.data.table(iris)
iris_levels = levels(iris_dt$Species)

tsk_iris = as_task_torch(iris_dt, target = "Species", id = "iris",
  output_dim = function(task) length(iris_levels),
  prediction_encoder = function(task, predict_tensor, predict_type) {
    idx = as.integer(predict_tensor$argmax(dim = 2L)$cpu())
    list(response = factor(idx, levels = seq_along(iris_levels), labels = iris_levels))
  }
)
```

The task says the network has three outputs and that the largest of them names the predicted class.
Whether those three numbers are logits trained with cross entropy, or one-hot regression targets trained with mean squared error, is the learner's decision:

```{r custom_task_type-018}
lrn_onehot = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = t_loss("mse"),
  target_batchgetter = function(data) {
    torch_tensor(1 * model.matrix(~ 0 + data[[1L]]), dtype = torch_float())
  },
  epochs = 30, batch_size = 32, opt.lr = 0.05
)

lrn_onehot$train(tsk_iris)
lrn_onehot$predict(tsk_iris)$score(msr_torch("acc", function(truth, response) mean(truth == response)))
```

Swapping in a learner with `loss = t_loss("cross_entropy")` and a `target_batchgetter` returning class indices trains the same task as ordinary multi-class classification, and both produce the same kind of prediction, because the task decides that.

## Tasks without a Target {#sec-unsupervised}

A `TaskTorch` is supervised or unsupervised depending only on whether you gave it target columns.
Nothing beyond that is assumed about the structure of the problem.

A task with no targets has no target element in its batches at all, so the loss is called as `loss(y_hat)`, with no second argument at all.

If the target of a batch is a function of its *input* -- an autoencoder reconstructing its input, a denoising or masked objective, contrastive pretraining -- then the learner's `target_batchgetter` may declare an `x` argument, which receives the feature tensors of the batch.
An autoencoder over the numeric features of `iris` is then:

```{r custom_task_type-019}
iris_scaled = as.data.table(scale(iris[, 1:4]))

tsk_ae = as_task_torch(iris_scaled, id = "iris_ae",
  output_dim = function(task) length(task$feature_names),
  prediction_encoder = function(task, predict_tensor, predict_type) {
    response = as.matrix(predict_tensor$cpu())
    colnames(response) = task$feature_names
    list(response = response)
  }
)
tsk_ae
```

Such a task has no `truth`, so its measure reads the ground truth from the task.
`msr_torch()` arranges that for any function that declares a `task` argument:

```{r custom_task_type-020}
msr_recon = msr_torch("reconstruction", function(task, prediction) {
  truth = as.matrix(task$data(rows = prediction$row_ids, cols = task$feature_names))
  mean((truth - prediction$response)^2)
}, range = c(0, Inf))
```

The network is an ordinary autoencoder, and the loss compares its output to the target that the batchgetter produced:

```{r custom_task_type-021}
nn_ae = nn_module("nn_ae",
  initialize = function(task) {
    d_in = length(task$feature_names)
    self$encoder = nn_sequential(nn_linear(d_in, 16), nn_relu(), nn_linear(16, 2))
    self$decoder = nn_sequential(nn_linear(2, 16), nn_relu(), nn_linear(16, d_in))
  },
  forward = function(input) self$decoder(self$encoder(input))
)

lrn_ae = lrn("torch.module",
  module_generator = nn_ae,
  ingress_tokens = list(input = ingress_num()),
  loss = t_loss("mse"),
  # a reconstruction objective: the target of a batch is the batch's own input
  target_batchgetter = function(data, x) x[[1L]],
  epochs = 100, batch_size = 32, opt.lr = 0.01
)

lrn_ae$train(tsk_ae)
lrn_ae$predict(tsk_ae)$score(msr_recon, task = tsk_ae)
```

The features were standardized to unit variance, so a mean squared reconstruction error well below `1` means that a two-dimensional latent space captures most of the structure of the four original features.

Resampling works here too, and the task is passed on automatically:

```{r custom_task_type-022}
resample(tsk_ae, lrn_ae, rsmp("cv", folds = 3))$aggregate(msr_recon)
```

The latent codes -- arguably the more interesting output of an autoencoder -- are available through the trained network:

```{r custom_task_type-023}
network = lrn_ae$network
network$eval()
x = torch_tensor(as.matrix(tsk_ae$data(rows = 1:5)), dtype = torch_float())
as.matrix(with_no_grad(network$encoder(x)))
```

## What You Give Up {#sec-limits}

All of this convenience comes from one decision: every `TaskTorch` has the same task type, `"torch"`.
`mlr3` therefore cannot tell one custom problem from another.

```{r custom_task_type-024}
tsk_weather$task_type
tsk_ae$task_type
```

The consequences are worth being explicit about:

* A learner built for one problem will accept a task from a completely different one. If the tensor shapes happen to line up, it trains, and you get a meaningless model rather than an error.
* The same holds for measures: `mlr3` will not stop you scoring a multi-label prediction with a reconstruction error.
* There is no dictionary entry for your problem. `lrn("torch.module")` and `po("torch_model")` are shared by everyone, so you cannot hand a colleague a learner id and have it mean your problem.

> **When to write a real task type instead**
>
> If your learning problem is going into a package that other people depend on, add a proper task type to `mlr3` rather than using `TaskTorch`: a `Task` class, a `Prediction` class with its `PredictionData` methods, a `Measure`, entries in `mlr_reflections`, and methods for the `output_dim_for()`, `get_target_batchgetter()` and `encode_prediction()` generics.
> That is what `mlr3proba` does for survival analysis.
> You get real type checking, dictionary entries, and channel typing in `mlr3pipelines` -- at a cost of a few hundred lines instead of five.

`TaskTorch` is for the stage before that: exploring a problem, running an experiment, or trying an architecture on something unusual without committing to an API.

## Conclusion {#sec-conclusion}

In this article we trained networks on two problems that `mlr3torch` has no built-in support for -- multi-label classification and an autoencoder -- without adding a task type to `mlr3`.

`as_task_torch()` builds a task from any number of target columns, or none at all. You state the output dimension and the prediction encoding on the task and the target tensor on the learner; nothing is inferred from the data.
Each of the three can be overridden on the task, and the target encoding can additionally be overridden on the learner.
`msr_torch()` turns a plain R function into a measure.
Everything that is a modelling decision -- the loss and the measure -- you state yourself, because those do not follow from the data.
