---
title: "Custom Learning Problems"
author: "Sebastian Fischer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Custom Learning Problems}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r custom_task_type-001, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lgr::get_logger("mlr3")$set_threshold("warn")
```

This article will show how to use {mlr3torch} for tasks that go beyond the classification and regression tasks that are supported out of the box.
This is possible via the generic `TaskTorch`, which allows you to use mlr3torch with arbitrary modeling problems.
The price of this flexibility is fewer compatibility checks and thus more responsibility on the user, see [Drawbacks of `TaskTorch`](#sec-drawbacks) for more information.
In this article we will first show how to use `TaskTorch` in a multi-label classification problem and then how it can be used for an unsupervised autoencoder.

## Multi-label Classification

As an example, we will train a learner on a multi-label classification problem which consists of $k$ binary prediction problems where multiple labels can be true.
Below, we generate some synthetic data for it and want to predict from the five `x` features whether it's sunny, warm, and windy.

```{r custom_task_type-002}
library(mlr3torch)
library(data.table)

set.seed(314)
n = 500
dat = data.table(x1 = rnorm(n), x2 = rnorm(n), x3 = rnorm(n), x4 = rnorm(n), x5 = rnorm(n))
dat[, `:=`(
  sunny = x1 + x2 > 0,
  warm = x2 * x3 > 0,
  windy = x4 - x5 > 0
)]
```

In order to create a `TaskTorch` for this modeling problem, we want to first specify:

1. The default method for converting torch predictions to an `mlr3::Prediction` object,
2. The default `mlr3::Measure` for scoring these predictions, and
3. How many output units a network for this task needs.

For our problem, we want to accept both response (class) predictions and probability predictions, and we assume that the network outputs per-class logits.
Any learner that we create for this task can also deviate from this default encoding.

```{r custom_task_type-003}
weather_encoder = function(task, network_output, predict_type) {
  prob = as.matrix(nnf_sigmoid(network_output)$cpu())
  colnames(prob) = task$target_names
  list(response = prob > 0.5, prob = if (predict_type == "prob") prob)
}
```

For more information on what a prediction may hold and how it is tabled, see `?PredictionTorch`.
As the default measure, we use the *Hamming loss* and construct it via `msr_torch`, see its help page for more information.

```{r custom_task_type-004}
msr_hamming = msr_torch("multilabel.hamming",
  function(truth, response) mean(as.matrix(truth) != response),
  range = c(0, 1), minimize = TRUE
)
msr_hamming
```

Below, we create the `TaskTorch`, where the third of these is the `output_dim` argument, a function of the task.

```{r custom_task_type-005}
tsk_weather = as_task_torch(dat, target = c("sunny", "warm", "windy"), id = "weather",
  output_dim = function(task) length(task$target_names),
  default_encoder = weather_encoder,
  default_measure = msr_hamming
)
tsk_weather

tsk_weather$truth(1:5)
tsk_weather$target_names
tsk_weather$feature_names
```

The `output_dim` field is used via the `output_dim_for()` generic:

```{r custom_task_type-006}
output_dim_for(tsk_weather)
```

Below, we define the architecture where the output dimension is obtained using `output_dim_for()`.

```{r custom_task_type-007}
nn_mlp = nn_module("nn_mlp",
  initialize = function(task, latent) {
    self$net = nn_sequential(
      nn_linear(length(task$feature_names), latent), nn_relu(),
      nn_linear(latent, latent), nn_relu(),
      nn_linear(latent, output_dim_for(task))
    )
  },
  # the argument name matches the name of the ingress token below
  forward = function(input) self$net(input)
)
```

To turn this into a learner, we additionally need a loss and a way to construct the batches.
For the features, we use a standard numeric encoder via `ingress_num()`, so we only specify the `target_batchgetter` to also encode the targets numerically.
The data argument of the batchgetter is `task$data(batch_ids, task$target_names)` and it must return a torch tensor.
For the loss, we use the binary cross entropy, built from the corresponding torch loss:

```{r custom_task_type-008}
loss_bce = as_torch_loss(nn_bce_with_logits_loss, id = "bce")
loss_bce

# the loss wants a float tensor of zeros and ones, so that is what the learner builds
weather_batchgetter = function(data) torch_tensor(as.matrix(data), dtype = torch_float())

lrn_weather = lrn("torch.module",
  module_generator = nn_mlp,
  ingress_tokens = list(input = ingress_num()),
  loss = loss_bce,
  target_batchgetter = weather_batchgetter,
  predict_types = c("response", "prob"),
  latent = 32,
  epochs = 5, batch_size = 32, opt.lr = 0.01
)
lrn_weather$predict_type = "prob"
```

Next, we train the learner on two thirds of the data and make predictions on the remaining observations:

```{r custom_task_type-009}
split = partition(tsk_weather)
lrn_weather$train(tsk_weather, split$train)
pred = lrn_weather$predict(tsk_weather, split$test)
pred
```

The prediction is scored like any other.
Note that a `mlr3::Measure` differs from a `TorchLoss`: it is computed from the encoded prediction, whereas the torch loss operates on the raw network output and the tensor-encoded target.

```{r custom_task_type-010}
pred$score(msr_hamming)
```

It is also possible to get the raw network outputs as a `lazy_tensor`:

```{r custom_task_type-011}
lrn_logits = lrn_weather$clone(deep = TRUE)
lrn_logits$predict_type = "lazy_tensor"

pred_logits = lrn_logits$predict(tsk_weather, row_ids = split$test[1:3])
pred_logits
materialize(pred_logits$lazy_tensor, rbind = TRUE)
```

Of course, we can also resample the learner.

```{r custom_task_type-012}
rr = resample(tsk_weather, lrn_weather, rsmp("cv", folds = 3))
rr$aggregate()
```

We can also construct the learner as a `Graph`.
Here, we have to specify the `target_batchgetter` in the `PipeOpTorchModel`.
Note that `nn("head")` also relies on `output_dim_for` to define the last layer.

```{r custom_task_type-013}
architecture = po("torch_ingress_num") %>>%
  nn("linear_1", out_features = 20) %>>%
  nn("relu_1") %>>%
  nn("linear_2", out_features = 20) %>>%
  nn("relu_2") %>>%
  nn("head") %>>%
  po("torch_loss", loss_bce) %>>%
  po("torch_optimizer", "adam", lr = 0.01) %>>%
  po("torch_model", batch_size = 32, epochs = 50,
    target_batchgetter = weather_batchgetter)

glrn_weather = as_learner(architecture)
glrn_weather$train(tsk_weather, row_ids = split$train)
glrn_weather$predict(tsk_weather, row_ids = split$test)$score(msr_hamming)
```

## A Simple Autoencoder {#sec-unsupervised}

A `TaskTorch` is supervised or unsupervised depending only on whether you gave it target columns.
Nothing beyond that is assumed about the structure of the problem.

A task with no targets has no target element in its batches at all, so the loss is called as `loss(y_hat)`, with no second argument.

If the target of a batch is a function of its *input*, as it is for an autoencoder reconstructing its input, a denoising or masked objective, or contrastive pretraining, then the learner's `target_batchgetter` may declare an `x` argument, which receives the feature tensors of the batch.
An autoencoder over the numeric features of `iris` is then:

```{r custom_task_type-014}
iris_scaled = as.data.table(scale(iris[, 1:4]))

tsk_ae = as_task_torch(iris_scaled, id = "iris_ae",
  output_dim = function(task) length(task$feature_names),
  default_encoder = function(task, network_output, predict_type) {
    response = as.matrix(network_output$cpu())
    colnames(response) = task$feature_names
    list(response = response)
  }
)
tsk_ae
```

Such a task has no `truth`, so its measure reads the ground truth from the task.
`msr_torch()` arranges that for any function that declares a `task` argument:

```{r custom_task_type-015}
msr_recon = msr_torch("reconstruction", function(task, prediction) {
  truth = as.matrix(task$data(rows = prediction$row_ids, cols = task$feature_names))
  mean((truth - prediction$response)^2)
}, range = c(0, Inf))
```

The network is an ordinary autoencoder, and the loss compares its output to the target that the batchgetter produced:

```{r custom_task_type-016}
nn_ae = nn_module("nn_ae",
  initialize = function(task) {
    d_in = length(task$feature_names)
    self$encoder = nn_sequential(
      nn_linear(d_in, 16),
      nn_relu(),
      nn_linear(16, 2)
    )
    self$decoder = nn_sequential(
      nn_linear(2, 16),
      nn_relu(),
      nn_linear(16, d_in)
    )
  },
  forward = function(input) self$decoder(self$encoder(input))
)

lrn_ae = lrn("torch.module",
  module_generator = nn_ae,
  ingress_tokens = list(input = ingress_num()),
  loss = t_loss("mse"),
  # a reconstruction objective: the target of a batch is the batch's own input
  target_batchgetter = function(data, x) x[[1L]],
  epochs = 100, batch_size = 32, opt.lr = 0.01
)

lrn_ae$train(tsk_ae)
lrn_ae$predict(tsk_ae)$score(msr_recon, task = tsk_ae)
```

### Predictions as Tensors {#sec-raw-predictions}

The encoder above exists to turn the network's output into something a measure can be written against.
When the output *is* what you want, the predict type `"lazy_tensor"` hands it over directly, and the task then needs no `default_encoder` at all:

```{r custom_task_type-017}
lrn_raw = lrn_ae$clone(deep = TRUE)
lrn_raw$predict_type = "lazy_tensor"

pred_raw = lrn_raw$predict(tsk_ae)
pred_raw
```

One element per observation, holding that observation's reconstruction.
A network with more than one head gives you one such column per head, named after the heads.
`materialize()` turns them back into a tensor:

```{r custom_task_type-018}
reconstruction = materialize(pred_raw$lazy_tensor, rbind = TRUE)
reconstruction$shape
```

It is the same network output the encoder saw, just not encoded into anything:

```{r custom_task_type-019}
encoded = lrn_ae$predict(tsk_ae)$response
all.equal(as.matrix(reconstruction$cpu()), unname(encoded), tolerance = 1e-6)
```

This is the one predict type whose predictions do **not** survive `saveRDS()`.
They hold `torch` tensors, which are external pointers: saving succeeds and the tensors are invalid the next time they are touched, in this session or another. A `ResampleResult` containing them keeps its row ids and scores but loses its tensors.
Keep the encoder for anything you want to store, and reach for `"lazy_tensor"` when you are working with the output there and then.

The latent codes, arguably the more interesting output of an autoencoder, are not a prediction at all: nothing can score them and there is no ground truth to compare them against.
They are a property of the network, and the network is right there in the trained learner.
The only thing you need from `mlr3torch` is the tensors that the network expects, which `learner$dataset()` builds the same way training and prediction build them:

```{r custom_task_type-020}
ds = lrn_ae$dataset(tsk_ae)
batch = ds$.getbatch(1:5)

network = lrn_ae$network
network$eval()
as.matrix(with_no_grad(network$encoder(batch$x$input))$cpu())
```

Building that tensor by hand would work for a task whose features are four numeric columns, and would quietly go wrong for one with a `factor`, a `lazy_tensor` or an image transformation, because the ingress tokens are what decide how a batch is built.
Reach for `learner$dataset()` whenever you want to run the network yourself: intermediate activations, gradients with respect to an input, anything that is not a prediction.
For larger tasks, hand the dataset to `torch::dataloader()` rather than asking for every row at once, and move the tensors to the learner's device yourself.

## Drawbacks of `TaskTorch` {#sec-drawbacks}

Now that we have shown the capabilities of `TaskTorch`, it is also important to highlight the drawbacks of the approach.
In {mlr3}, objects are annotated with meta-information that describes their properties and capabilities.
For example, training a classification learner on a regression task raises a clear error message:

```{r custom_task_type-021, error = TRUE}
lrn("classif.featureless")$train(tsk("mtcars"))
```

This is because they have different `task_type`s.
Because the `"torch"` task type can represent an arbitrary task type (that is what it is designed for), we can't distinguish between the multi-label classification task and the reconstruction problem.
E.g., training the autoencoder on the weather data is simply undefined behavior and we cannot guarantee anything.
The same also holds true for measures.
As such, the `TaskTorch` is a *quick and dirty* way to approach a modeling problem.
The alternative is to properly create a new task type, which requires implementing sub-classes (such as `TaskRegr`, `MeasureRegr`, etc.) and registering the new task type with {mlr3}.
For a simple one-off modeling problem, the effort involved might not be worth it.