---
title: "Writing your own PipeOpTorch"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing your own PipeOpTorch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(mlr3torch)
```

The {mlr3torch} package already provides `PipeOpTorch` objects for many common neural network layers.
However, sooner or later you will want to add your own custom layer.
This article will explain you how to do so.
We assume that you already understand how to use them, which is covered in [Defining an Architecture](pipeop_torch.html).
The *Inheriting* section of `PipeOpTorch` also explains how to create new `PipeOpTorch` classes in more detail, whereas this article is a more practically oriented guide.

A `PipeOpTorch` is a descriptor of an `nn_module_generator`: it does not contain a module, it says
how to build one once the shape of its input is known.
We will first create a custom module generator and then turn it into a `PipeOpTorch`.
As an example we re-implement `torch::nn_linear()`.

## A module generator

A module generator is created with `torch::nn_module()` and it requires implementing two methods:

1. `$initialize()`, which constructs the instance, and 
2. `$forward()`, which runs the constructed layer on some tensor inputs.

```{r}
nn_my_linear = nn_module("nn_my_linear",
  initialize = function(in_features, out_features) {
    self$weight = nn_parameter(torch_randn(out_features, in_features) / sqrt(in_features))
    self$bias = nn_parameter(torch_zeros(out_features))
  },
  forward = function(input) {
    nnf_linear(input, self$weight, self$bias)
  }
)
```

Calling the generator builds a module instance (via `$initialize()` ) and calling the resulting module on an input tensor runs its `$forward()` method:

```{r}
layer = nn_my_linear(in_features = 4, out_features = 2)
layer(torch_randn(3, 4))
```

## The same layer as a `PipeOpTorch`

For `nn_my_linear`, the `in_features` need to be manually specified.
However, they are an *auxiliary* parameter, in the sense that they can be inferred from the input shape that the layer will receive during runtime.
The core idea of a `PipeOpTorch` is that it does not require specifying such auxiliary parameters but that they are automatically determined.
To enable this, a `PipeOpTorch` needs to define how to infer these auxiliary parameters and how to compute the output shapes for the specified input shapes and hyperparameters.
In the code below, the `$initialize()` method of the module generator takes in the `shapes_in` and  computes the `in_features` from their last dimension.

```{r}
PipeOpTorchMyLinear = pipeop_torch("nn_my_linear",
  initialize = function(shapes_in, out_features) {
    in_features = tail(shapes_in[[1L]], 1L)
    self$weight = nn_parameter(torch_randn(out_features, in_features) / sqrt(in_features))
    self$bias = nn_parameter(torch_zeros(out_features))
  },
  forward = function(input) {
    nnf_linear(input, self$weight, self$bias)
  },
  shapes_out = function(shapes_in, param_vals) {
    list(c(head(shapes_in[[1L]], -1L), param_vals$out_features))
  }
)
```

`out_features` is the only hyperparameter, and the operator can be used like any other:

```{r}
po_linear = PipeOpTorchMyLinear$new(param_vals = list(out_features = 2))
po_linear$param_set$ids()
po_linear$shapes_out(list(c(NA, 4)))

md = (po("torch_ingress_num") %>>% po_linear)$train(tsk("iris"))[[1L]]
model_descriptor_to_module(md)
```

The rest of this article explains the pieces: what the operator has to know, how to write one whose
module already exists, and how to write the class by hand instead of generating it.

## What a `PipeOpTorch` has to know

A network in `mlr3torch` is assembled without any data flowing through it.
Instead of tensors, the `Graph` passes around a `ModelDescriptor` that carries the *shape* of the
tensors that will later be there.
That is what makes `po("nn_linear", out_features = 10)` possible: you say how wide the layer should
be, and `in_features` follows from whatever is connected to its input.

A `PipeOpTorch` therefore answers two questions that the module itself does not:

1. **What does the module need to know about its input?** `in_features` of `nn_linear()` is not
   something the user sets, it is read off the incoming shape.
2. **What shape does the module produce?** The next layer needs it, for the same reason.

Everything else -- building the module, adding it to the network, connecting the edges -- is
inherited.

## The short way: `pipeop_torch()`

Take another layer, one that scales each feature by a learned factor.
Written with `pipeop_torch()`, it looks exactly like `nn_module()` would, with two additions:
`initialize()` takes `shapes_in` -- question 1 -- and `shapes_out` answers question 2.

```{r}
PipeOpTorchCustomScale = pipeop_torch("nn_custom_scale",
  initialize = function(shapes_in, init = 1) {
    self$weight = nn_parameter(torch_full(tail(shapes_in[[1L]], 1L), init))
  },
  forward = function(input) input * self$weight,
  shapes_out = function(shapes_in) shapes_in
)
```

The remaining arguments of `initialize()` become the hyperparameters of the operator: `init` is
something the user sets, the number of features is not.

```{r}
po_custom_scale = PipeOpTorchCustomScale$new()
po_custom_scale$param_set$ids()
po_custom_scale$shapes_out(list(c(NA, 4)))
```

Note the `NA`: the batch dimension is unknown while the network is being built, and any other
dimension may be too -- the sequence length of a transformer, or the size of an image.
Both `initialize()` and `shapes_out()` see those `NA`s and have to cope with them, either by
carrying them through, as here, or by saying which ones they need.
The helpers `assert_known_dims()`, `assert_ndim()` and friends state which dimensions you actually
need, and report the offending `PipeOp` and shape when one of them is unknown.
Be permissive with the rest: rejecting a shape that would have worked is worse than a runtime error
later.

That is a usable operator:

```{r}
graph = po("torch_ingress_num") %>>% po_custom_scale %>>% po("nn_head")
md = graph$train(tsk("iris"))[[1L]]
network = model_descriptor_to_module(md)
network
```

The module was built with four features, taken from the task:

```{r}
as.numeric(md$graph$pipeops$nn_custom_scale$module$weight)
```

Besides `shapes_in`, `initialize()` can take the `task`, and `shapes_out()` can take the
`param_vals` and the `task` as well -- each is passed only if the function declares it.
That is how `PipeOpTorchMyLinear` above states its output shape: the width of the layer is a
hyperparameter, so `shapes_out()` reads it from `param_vals`.
Since `out_features` has no default there, it is a *required* hyperparameter and the operator
refuses to be used without it.

```{r, error = TRUE}
PipeOpTorchMyLinear$new()$shapes_out(list(c(NA, 4)))
```

A module that already exists does not have to be rewritten like this: call its generator from
`initialize()` -- `self$linear = nn_linear(tail(shapes_in[[1L]], 1L), out_features)` -- and delegate
to it in `forward()`.

## Several channels

A module whose `forward()` takes more than one tensor becomes a `PipeOp` with more than one input
channel -- the channels are named after the arguments, and `shapes_in` is named after the channels.
Returning a list makes several output channels, which have to be declared with `out_channels`:

```{r}
PipeOpTorchSplit = pipeop_torch("nn_split",
  initialize = function(shapes_in, d_out1, d_out2) {
    self$linear1 = nn_linear(tail(shapes_in$input, 1L), d_out1)
    self$linear2 = nn_linear(tail(shapes_in$input, 1L), d_out2)
  },
  forward = function(input) {
    list(self$linear1(input), self$linear2(input))
  },
  out_channels = 2,
  shapes_out = function(shapes_in, param_vals) {
    list(c(head(shapes_in$input, -1L), param_vals$d_out1),
      c(head(shapes_in$input, -1L), param_vals$d_out2))
  }
)

po_split = PipeOpTorchSplit$new(param_vals = list(d_out1 = 10, d_out2 = 20))
po_split$shapes_out(list(c(NA, 4)))
```

`in_channels` and `out_channels` also take a count instead of names -- `in_channels = 0` for a
module whose `forward()` takes `...`, i.e. a *vararg* channel.

## Making the operator available as `nn("custom_scale")`

So far the class is only the value `pipeop_torch()` returned; `po()` and `nn()` know nothing about
it, because they look their arguments up in the `mlr_pipeops` dictionary.
Adding it there is one line:

```{r}
mlr_pipeops$add("nn_custom_scale", PipeOpTorchCustomScale)
```

`nn()` prefixes its key with `nn_` to find the entry and uses the unprefixed key as the id, so the
new layer is now written exactly like the ones that come with {mlr3torch}:

```{r}
po("torch_ingress_num") %>>% nn("custom_scale", init = 2) %>>% nn("head")
```

When you ship an operator in a package, that call belongs in the package's `.onLoad()`, so that the
dictionary is filled when the package is loaded.

## The long way: inheriting from `PipeOpTorch`

`pipeop_torch()` writes an R6 class for you; when you want to ship an operator in a package, or
need control over its construction, write that class yourself.
The module is then an ordinary `torch` module that knows nothing about shapes,

```{r}
nn_custom_scale = nn_module("nn_custom_scale",
  initialize = function(n_features, init = 1) {
    self$weight = nn_parameter(torch_full(n_features, init))
  },
  forward = function(input) input * self$weight
)
```

and the two questions are the private methods `.shape_dependent_params()` and `.shapes_out()`:

```{r}
PipeOpTorchCustomScale = R6::R6Class("PipeOpTorchCustomScale",
  inherit = PipeOpTorch,
  public = list(
    initialize = function(id = "nn_custom_scale", param_vals = list()) {
      param_set = paradox::ps(
        init = paradox::p_dbl(default = 1, tags = "train")
      )
      super$initialize(
        id = id,
        param_vals = param_vals,
        param_set = param_set,
        module_generator = nn_custom_scale
      )
    }
  ),
  private = list(
    .shape_dependent_params = function(shapes_in, param_vals, task) {
      c(param_vals, list(n_features = tail(shapes_in[[1L]], 1L)))
    },
    .shapes_out = function(shapes_in, param_vals, task) shapes_in
  )
)

PipeOpTorchCustomScale$new()$shapes_out(list(c(NA, 4)))
```

`.shape_dependent_params()` returns *all* arguments the module is built with, i.e. the user's
`param_vals` plus the inferred ones.
If the module cannot be built by calling the generator with a list of arguments at all, overload
`.make_module(shapes_in, param_vals, task)` instead and ignore both.

A class written this way is added to the dictionary just like a generated one, see
[above](#making-the-operator-available-as-nncustom_scale).

## Registering the operator with a learner

Nothing else is needed: a `Graph` that contains the new operator becomes a `Learner` the usual way.

```{r}
graph = po("torch_ingress_num") %>>%
  PipeOpTorchCustomScale$new() %>>%
  po("nn_head") %>>%
  po("torch_loss", t_loss("cross_entropy")) %>>%
  po("torch_optimizer", t_opt("adam")) %>>%
  po("torch_model_classif", batch_size = 32, epochs = 1)

learner = as_learner(graph)
learner$id = "scaled_mlp"
learner
```

## Checking your work

Two things are worth testing, and both are cheap:

* `$shapes_out()` agrees with what the module actually produces, for shapes with and without `NA`s.
  `mlr3torch`'s own test suite sweeps every combination of unknown dimensions for every operator it
  ships; the helper that does it lives in `tests/testthat/helper_shape_inference.R`.
* the operator survives a round-trip through a `Graph`, i.e. `model_descriptor_to_module()` returns
  a network that runs on a tensor of the announced shape.

```{r}
po_custom_scale = PipeOpTorchCustomScale$new()
inferred = po_custom_scale$shapes_out(list(c(NA, 4)))[[1L]]

network = model_descriptor_to_module(
  (po("torch_ingress_num") %>>% po_custom_scale)$train(tsk("iris"))[[1L]]
)
observed = dim(with_no_grad(network(torch_ingress_num.input = torch_randn(2, 4))))

inferred
observed
```
