---
title: "Writing your own PipeOpTorch"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing your own PipeOpTorch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(mlr3torch)
```

`mlr3torch` ships a `PipeOp` for most of the layers that `torch` provides, but sooner or later you
will want a layer of your own in a `Graph`.
This article shows how to get there, from the shortest path to the full R6 class.

If you have not built a network from `PipeOp`s before, read
[Defining an Architecture](pipeop_torch.html) first; for what a `ModelDescriptor` is and how a
`Graph` of `PipeOp`s becomes a `nn_module`, see [Networks as Graphs](internals_pipeop_torch.html).

## What a `PipeOpTorch` has to know

A network in `mlr3torch` is assembled without any data flowing through it.
Instead of tensors, the `Graph` passes around a `ModelDescriptor` that carries the *shape* of the
tensors that will later be there.
That is what makes `po("nn_linear", out_features = 10)` possible: you say how wide the layer should
be, and `in_features` follows from whatever is connected to its input.

A `PipeOpTorch` therefore answers two questions that the module itself does not:

1. **Which of the module's arguments follow from the input shape?** `in_features` of
   `nn_linear()` is not something the user sets, it is read off the incoming shape.
2. **What shape does the module produce?** The next layer needs it, for the same reason.

Everything else -- building the module, adding it to the network, connecting the edges -- is
inherited.

## The short way: `pipeop_torch()`

Take a layer that scales each feature by a learned factor:

```{r}
nn_scale = nn_module("nn_scale",
  initialize = function(n_features, init = 1) {
    self$weight = nn_parameter(torch_full(n_features, init))
  },
  forward = function(input) input * self$weight
)
```

`n_features` is question 1: the user cannot know it, it is the last dimension of the input.
`pipeop_torch()` takes it as an `auxiliary` argument -- a function of the input shapes -- and turns
everything else into a hyperparameter:

```{r}
PipeOpTorchScale = pipeop_torch("nn_scale", nn_scale,
  auxiliary = list(
    n_features = function(shapes_in, param_vals, task) tail(shapes_in[[1L]], 1L)
  )
)

po_scale = PipeOpTorchScale$new()
po_scale$param_set$ids()
```

Question 2 is answered by default: `pipeop_torch()` builds the module and traces it on the
`"meta"` device, which computes shapes without allocating any memory.

```{r}
po_scale$shapes_out(list(c(NA, 4)))
```

Note the `NA`: the batch dimension is unknown while the network is being built, and any other
dimension may be too -- the sequence length of a transformer, or the size of an image.
Shape inference has to carry those through rather than reject them.

That is a usable operator:

```{r}
graph = po("torch_ingress_num") %>>% po_scale %>>% po("nn_head")
md = graph$train(tsk("iris"))[[1L]]
network = model_descriptor_to_module(md)
network
```

The module was built with `n_features = 4`, taken from the task:

```{r}
as.numeric(md$graph$pipeops$nn_scale$module$weight)
```

If a module needs nothing from the input shape at all, there is nothing left to configure and
`as_pipeop()` is enough:

```{r}
nn_square = nn_module("nn_square", forward = function(input) input^2)
as_pipeop(nn_square)$shapes_out(list(c(NA, 4)))
```

### When to write `shapes_out` yourself

Tracing is convenient but it builds the module, and it is only correct when the output shape
depends on the shapes of the inputs rather than on the values in them.
Pass a function to say it directly:

```{r}
PipeOpTorchScale = pipeop_torch("nn_scale", nn_scale,
  auxiliary = list(
    n_features = function(shapes_in, param_vals, task) tail(shapes_in[[1L]], 1L)
  ),
  shapes_out = function(shapes_in, param_vals, task) shapes_in
)
```

For an operator that genuinely does not change the shape -- activation functions, dropout --
`shapes_out = NULL` says exactly that.

Whatever you write must cope with `NA` in any dimension.
The helpers `assert_known_dims()`, `assert_ndim()` and friends state which dimensions you actually
need, and report the offending `PipeOp` and shape when one of them is unknown.
Be permissive with the rest: rejecting a shape that would have worked is worse than a runtime error
later.

## Several channels

A module whose `forward()` takes more than one tensor becomes a `PipeOp` with more than one input
channel -- the channels are named after the arguments.
Returning a named list makes several output channels, which have to be declared with `outname`:

```{r}
nn_split = nn_module("nn_split",
  initialize = function(d_in, d_out1, d_out2) {
    self$linear1 = nn_linear(d_in, d_out1)
    self$linear2 = nn_linear(d_in, d_out2)
  },
  forward = function(input) {
    list(output1 = self$linear1(input), output2 = self$linear2(input))
  }
)

PipeOpTorchSplit = pipeop_torch("nn_split", nn_split,
  outname = c("output1", "output2"),
  auxiliary = list(d_in = function(shapes_in, param_vals, task) tail(shapes_in[[1L]], 1L))
)

po_split = PipeOpTorchSplit$new(param_vals = list(d_out1 = 10, d_out2 = 20))
po_split$shapes_out(list(c(NA, 4)))
```

## The long way: inheriting from `PipeOpTorch`

`pipeop_torch()` writes an R6 class for you; when you want to ship an operator in a package, or
need control over its construction, write that class yourself.
The two questions are now the private methods `.shape_dependent_params()` and `.shapes_out()`:

```{r}
PipeOpTorchScale = R6::R6Class("PipeOpTorchScale",
  inherit = PipeOpTorch,
  public = list(
    initialize = function(id = "nn_scale", param_vals = list()) {
      param_set = paradox::ps(
        init = paradox::p_dbl(default = 1, tags = "train")
      )
      super$initialize(
        id = id,
        param_vals = param_vals,
        param_set = param_set,
        module_generator = nn_scale
      )
    }
  ),
  private = list(
    .shape_dependent_params = function(shapes_in, param_vals, task) {
      c(param_vals, list(n_features = tail(shapes_in[[1L]], 1L)))
    },
    .shapes_out = function(shapes_in, param_vals, task) shapes_in
  )
)

PipeOpTorchScale$new()$shapes_out(list(c(NA, 4)))
```

`.shape_dependent_params()` returns *all* arguments the module is built with, i.e. the user's
`param_vals` plus the inferred ones.
If the module cannot be built by calling the generator with a list of arguments at all, overload
`.make_module(shapes_in, param_vals, task)` instead and ignore both.

To make the operator available as `po("nn_scale")`, add it to the `PipeOp` dictionary, typically in
your package's `.onLoad()`:

```{r, eval = FALSE}
mlr3pipelines::mlr_pipeops$add("nn_scale", PipeOpTorchScale)
```

## Registering the operator with a learner

Nothing else is needed: a `Graph` that contains the new operator becomes a `Learner` the usual way.

```{r}
graph = po("torch_ingress_num") %>>%
  PipeOpTorchScale$new() %>>%
  po("nn_head") %>>%
  po("torch_loss", t_loss("cross_entropy")) %>>%
  po("torch_optimizer", t_opt("adam")) %>>%
  po("torch_model_classif", batch_size = 32, epochs = 1)

learner = as_learner(graph)
learner$id = "scaled_mlp"
learner
```

## Checking your work

Two things are worth testing, and both are cheap:

* `$shapes_out()` agrees with what the module actually produces, for shapes with and without `NA`s.
  `mlr3torch`'s own test suite sweeps every combination of unknown dimensions for every operator it
  ships; the helper that does it lives in `tests/testthat/helper_shape_inference.R`.
* the operator survives a round-trip through a `Graph`, i.e. `model_descriptor_to_module()` returns
  a network that runs on a tensor of the announced shape.

```{r}
po_scale = PipeOpTorchScale$new()
inferred = po_scale$shapes_out(list(c(NA, 4)))[[1L]]

network = model_descriptor_to_module(
  (po("torch_ingress_num") %>>% po_scale)$train(tsk("iris"))[[1L]]
)
observed = dim(with_no_grad(network(torch_ingress_num.input = torch_randn(2, 4))))

inferred
observed
```
