---
title: "Image Classification"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In the *Get Started* vignette, we have already explained how to train a simple neural network on tabular data.
In this article we will briefly cover how to work with image data.
In `mlr3torch`, image data is represented using the `mlr3torch::imageuri` class.
It is essentially a character vector, containing paths to images on the file system.
When listing the available task feature types (after loading the `mlr3torch` package), we can see that this class is available.

```{r}
library(mlr3torch)
mlr_reflections$task_feature_types
```

Creating a vector (in this case of length 1) is as simple as passing the image paths to the `imageuri` function.

```{r}
image_vec = imageuri("/path/to/your/image")
```

For the processing of images, `mlr3torch` relies mostly on the functionality provided by [`torchvision`](https://github.com/mlverse/torchvision).
As an example task, we will use the "tiny imagenet" dataset, which is a subset of the [ImageNet](http://www.image-net.org/) dataset.
It consists of 200 classes with 500 training images each.
The goal is to predict the class of an image from the pixels.
For more information you can access the tasks's help page.

```{r setup}
set.seed(314)
library(mlr3torch)

tsk_tiny = tsk("tiny_imagenet")
tsk_tiny
```

The first time this task is accessed, the data is downloaded from the internet.
In order to download the dataset only once, you can set the `mlr3torch.cache` option to either `TRUE` or a specific path to be used as the cache folder.

```{r, eval = FALSE}
options(mlr3torch.cache = TRUE)
```

Below, we print the path to the first image as follows:

```{r}
tsk_tiny$data(1, "image")
```

To work with such data, we need to use learners that have `"imageuri"` as part of their supported feature types.
One such learner is the famous AlexNet classification network, which sparked the "Deep Learning revolution" in 2012.

```{r}
alexnet = lrn("classif.alexnet")
alexnet
```

We can now train this learner like any other learner on the task at hand, while `mlr3torch` internally creates a dataloader from the image paths.
For computational reasons, we cannot demonstrate the actual training of the learner in this article.
