---
title: "Lazy Tensor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lazy Tensor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(ggplot2)
theme_set(theme_bw())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the `lazy_tensor` class, which is a vector type that can be used to lazily represent torch tensors of arbitary dimensions.
Among other things, it allows `mlr3torch` to work with images, which we will illustrate using the predefined MNIST task, which has one feature `image` of class `"lazy_tensor"`.
The images display the digits 0, ..., 9, and the goal is to classify them correctly.

```{r, message = FALSE}
library(mlr3torch)

mnist = tsk("mnist")
mnist
```

The name `"lazy_tensor"` stems from the fact, that the tensors are not necessarily stored in memory, as this is often impossible when working with large image datasets.  
Therefore, we can easily access the data without any expensive data-loading.
We see that the data contains one column *label*, which is the target variable, and an `image` which is the input feature.

```{r}
mnist$head()
```

If we wanted to obtain the actual tensors representing the images, we can do so by calling `materialize()`, which will return a list of `torch_tensor`s, not necessarily all with the same shape.
Here, we only show a slice of the tensor for readability.


```{r}
lt = mnist$data(cols = "image")[[1L]]
materialize(lt[1])[[1]][1, 12:16, 12:16]
```

If all elements have the same shape as is the case here, we could also obtain a single `torch_tensor` by specifying `rbind = TRUE`.

In order to train a `Learner` on a `Task` containing `lazy_tensor` columns it must support the `lazy_tensor` feature type, as is the case for the multi layer perceptron:

```{r}
mlp = lrn("classif.mlp",
  neurons = c(100, 100),
  epochs = 10, batch_size = 32
)
mlp
```

However, because `lazy_tensor`s also have a specific shape, we also must ensure that the shape of the `lazy_tensor` matches the expected input shape of the learner.
The multi layer perceptron expects a 2d-tensor where the first dimension is the batch dimension.
But above we have seen that this is not the case for MNIST, where each element has shape `(1, 28, 28)`.
Therefore, we need to flatten the `lazy_tensor`, which we here do using `po("trafo_reshape")`:

```{r}
reshaper = po("trafo_reshape", shape = c(-1, 28 * 28))
mnist_flat = reshaper$train(list(mnist))[[1L]]
mnist_flat$head()
```

Note that this does not *actually* reshape all the tensors, this will again only happen once `materialize()` is called.

We can now proceed to train the a simple multi-layer perceptron on the flattened mnist task:

```{r, eval = FALSE}
mlp = lrn("classif.mlp",
  neurons = c(100, 100),
  epochs = 10, batch_size = 32
)
mlp$train(mnist_flat)
```

```{r, include = FALSE, echo = FALSE, eval = TRUE}
# ensure that code above is working
mlp$param_set$values$epochs = 1L
mlp$train(mnist_flat$clone()$filter(1:10))
```

## Creating a Lazy Tensor

Every `lazy_tensor` is built on top of a `torch::dataset`, so we here assume that you are familiar with it.
For more information on how to create `torch::dataset`s, we recommend reading the [torch package documentation](https://torch.mlverse.org/).
The only additional restriction that we impose on the dataset is that it must have a `.getitem` or `.getbatch` method that returns a list of named tensors.

As an example, we will create a `lazy_tensor` of length 1000, whose elements are drawn from a uniform distribution over $[0, 1]$.
While the data is stored in-memory in this example, this is not necessary and the `$.getitem()` method can e.g. load images from disk.

```{r}
mydata = dataset(
  initialize = function() {
    self$x = runif(1000, -1, 1)
  },
  .getbatch = function(i) list(x = torch_tensor(self$x[i])$unsqueeze(2)),
  .length = function() 1000
)()
```

In order to create a `lazy_tensor` from `mydata`, we have to annotate the returned shapes of the dataset by passing a named list to `dataset_shapes`.
The first dimension must be `NA`, as it is the batch dimension.
We can also set a shape to `NULL` to indicate that it is unknown, i.e. it varies between elements.

```{r}
lt = as_lazy_tensor(mydata, dataset_shapes = list(x = c(NA, 1)))
lt[1:5]
```

Note that in this case, because we implemented the `.getbatch` method, we could have even omitted specifying the `dataset_shapes` as they could have been auto-inferred. 

We can convert this vector to a `torch_tensor` just like before

```{r}
materialize(lt[1:5], rbind = TRUE)
```

Because we added no preprocessing, this is the same as calling the `$.getbatch()` method on `mydata` and selecting the element `x`.

```{r}
torch_equal(
  materialize(lt[1], rbind = TRUE),
  mydata$.getbatch(1)$x
)
```

We continue with creating an example task from `lt`, where the relationship between the `x` and `y` variable is polynomial.
Note that the target variable, both for classification and regression, cannot be a `lazy_tensor`, but must be a `factor` or `numeric`.

```{r, message = FALSE}
library(data.table)
x = mydata$x
y = 0.2 + 0.1 * x - 0.1 * x^2 - 0.3 * x^3 + 0.5 * x^4 + 0.5 * x^7 + 0.6 * x^11 + rnorm(length(mydata)) * 0.1
dt = data.table(y = y, x = lt)
task_poly = as_task_regr(dt, target = "y", id = "poly")
task_poly
```

Below, we plot the data:

```{r, fig.align = "center"}
library(ggplot2)
ggplot(data = data.frame(x = x, y = y)) +
  geom_point(aes(x = x, y = y), alpha = 0.5)
```

In the next section, we will create a custom `PipeOp` to fit a polynomial regression model.

## Custom Preprocessing

In order to create a custom preprocessing operator for a lazy tensor, we have to create a new `PipeOp` class.
To make this as convenient as possible, `mlr3torch` offters a `pipeop_preproc_torch()` function that we recommend using for this purpose.
Its most important arguments are:

* `id` - Used as the default identifier of the `PipeOp`
* `fn` - The preprocessing function. The first argument is assumed to be the `torch_tensor` and the remaining arguments will be part of the `PipeOp`'s parameter set. 
* `shapes_out` - A function that returns the shapes of the output tensors given the input shapes.
  This can also be set to `NULL` for an unknown shape or to `"infer"` for auto-inference, see `?pipeop_preproc_torch` for more information.

Below, we create a `PipeOp`, that transforms a vector `x` into a matrix $(x^{d_1} ..., x^{d_n})$, where $d$ is the `degrees` parameter of the `PipeOp`.

```{r}
PipeOpPreprocTorchPoly = pipeop_preproc_torch("poly",
  fn = function(x, degrees) {
    torch_cat(lapply(degrees, function(d) torch_pow(x, d)), dim = 2L)
  },
  shapes_out = "infer"
)
```

We can now create a new instance of this `PipeOp` by calling `$new()`, and we set the parameter `degrees` to those degrees that were used when simulating the data above. 
Further, we set the parameter `stages`, that is always available, to `"both"`, which means that the preprocessing is applied during training and prediction.
For data augmentation, for example, this can be set to `"train"`.

```{r}
po_poly = PipeOpPreprocTorchPoly$new()

po_poly$param_set$set_values(
  degrees = c(0, 1, 2, 3, 4, 7, 11),
  stages = "both"
)
```

To create our polynomial regression learner, we combine the polynomial preprocessor with a `lrn("regr.mlp")` with no hidden layer and train the learner on the task.

```{r}
lrn_poly = as_learner(
  po_poly %>>% lrn("regr.mlp", batch_size = 256, epochs = 100,
  neurons = integer(0))
)

lrn_poly$train(task_poly)
pred = lrn_poly$predict(task_poly)
```

Below, we visualize the predictions and see that the model captured the non-linear relationship reasonably:

```{r, fig.align = "center"}
dt = melt(data.table(
  truth = pred$truth,
  response = pred$response,
  x = x),
  id.vars = "x", measure.vars = c("truth", "response")
)
dt$variable = factor(dt$variable, levels = c("truth", "response"))

ggplot(data = dt) +
  geom_point(aes(x = x, y = value, color = variable))
```

In the next section, we will briefly cover the implementation details of the `lazy_tensor`, which is not necessary to work with the data-type, so feel free to skip this part.

## Digging Into Internals

Internally, the `lazy_tensor` vector uses the `DataDescriptor` class to represent the (possibly) preprocessed data.
It is very similar to the `ModelDescriptor` class that is used to build up neural nerworks using `PipeOpTorch` objects.
The `DataDescriptor` stores a `torch::dataset`, an `mlr3pipelines::Graph` and some metadata.


```{r}
desc = DataDescriptor$new(
  dataset = mydata,
  dataset_shapes = list(x = c(NA, 1))
)
```

Per default, the preprocessing graph contains only a single `PipOpNop` that does nothing.

```{r}
desc
```

The printed output of the data descriptor informs us about:

* The number of `PipeOp`s contained in the preprocessing graph
* The output shapes of the dataset
* The input map, i.e. how the data is passed to the preprocessing graph, which is important when there are multiple inputs
* The `pointer`, which points to a specific channel of an output `PipeOp`.
  The output of this channel is the tensor represented by the `DataDescriptor`.
* The `shape`, which is the shape of the tensor at posititon `pointer`


A lazy tensor can be constructed from an integer vector and a `DataDescriptor`.
The integer vector specifies which element of the `DataDescriptor` the `lazy_tensor` contains.
Below, the first two elements of the `lazy_tensor` vector represent the same element of the `DataDescriptor`, while the third element represents a different element.
Note that all indices refer to the same `DataDescriptor`.

```{r}
lt = lazy_tensor(desc, ids = c(1, 1, 2))
materialize(lt, rbind = TRUE)
```

Internally, the lazy tensor is represented as a list of lists, containing an id and a `DataDescriptor`
Currently, there can only be a single `DataDescriptor` in a `lazy_tensor` vector.
The first element in the `lazy_tensor` then represents the first element from the `DataDescriptor`:

```{r}
unclass(lt[[1]])
```


What happens during `materialize(lt[1])` is the following:

```{r}
# get index and data descriptor
desc = lt[[1]][[2]]
id = lt[[1]][[1]]

# retrieve the batch <id> from the datast
dataset_output = desc$dataset$.getbatch(id)

# batch is reorganized according to the input map
graph_input = dataset_output[desc$input_map]
names(graph_input) = names(desc$graph$input$name)

# the reorganized batch is fed into the preprocessing graph
graph_output = desc$graph$train(graph_input, single_input = FALSE)

# the output pointed to by the pointer is returned
tensor = graph_output[[paste0(desc$pointer, collapse = ".")]]
tensor
```
Preprocessing a `lazy_tensor` vector adds new `PipeOp`s to the preprocessing graph and updates the metainformation like the pointer and output shape.
To show this, we create a simple example task, using the `lt` vector as a feature.

```{r}
taskin = as_task_regr(data.table(x = lt, y = 1:3), target = "y")

taskout = po_poly$train(list(taskin))[[1L]]

lt_out = taskout$data(cols = "x")$x

descout = lt_out[[1]][[2]]

descout

descout$graph
```
