---
title: "Defining an Architecture"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Defining an Architecture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we will show two ways to define a neural network architecture in `mlr3torch`.
We start with the direct route, where an existing `torch` module is turned into an `mlr3::Learner`.
Afterwards, we show how to build architectures as `mlr3pipelines::Graph`s, which allows to infer shapes from the task and to tune the architecture itself.

```{r, message = FALSE}
library(mlr3torch)
```

## From a `torch` Module to a `Learner`

If you already have a `torch::nn_module`, or want to write one yourself, `lrn("classif.module")` (or `lrn("regr.module")` for regression) wraps it into an `mlr3::Learner`.
This gives you the training loop, resampling and tuning infrastructure of `mlr3torch` without having to express the architecture as a `Graph`.

The module generator needs to take `task` as an argument, so that dimensions that depend on the data -- such as the number of features or classes -- can be inferred during `$train()`.
Its remaining arguments become hyperparameters of the resulting `Learner`.

```{r}
nn_one_layer = nn_module("nn_one_layer",
  initialize = function(task, size_hidden) {
    self$first = nn_linear(task$n_features, size_hidden)
    self$second = nn_linear(size_hidden, output_dim_for(task))
  },
  # the argument x corresponds to the ingress token x below
  forward = function(x) {
    self$second(nnf_relu(self$first(x)))
  }
)
```

The second ingredient are the `ingress_tokens`, which specify how the features of the task are converted into the tensors that are passed to the module's `forward()` method.
Their names must match the arguments of `forward()`, so here `ingress_num()` collects all numeric features into a single tensor `x`.

```{r}
module_learner = lrn("classif.module",
  module_generator = nn_one_layer,
  ingress_tokens = list(x = ingress_num()),
  size_hidden = 20,
  epochs = 10,
  batch_size = 16
)
module_learner$train(tsk("iris"))
module_learner$network
```

Note that the parameters of the module generator are by default inferred as untyped (`ParamUty`).
To make them tunable, pass an explicit `param_set` to `lrn("classif.module")` that describes them, e.g. `paradox::ps(size_hidden = paradox::p_int(1L, tags = "train"))`.

## Building an Architecture as a `Graph`

Writing the module directly gives you full control, but the architecture is then a black box: its layer sizes are fixed in R code and cannot be tuned, and the input dimensions have to be derived from the task by hand.
Building the network as a `Graph` instead addresses both points.
To show this, we create a simple CNN for the tiny-imagenet task, which is a subset of the well-known Imagenet benchmark.

```{r}
imagenet = tsk("tiny_imagenet")
imagenet
```

The central ingredients for creating such graphs are `PipeOpTorch` operators.

To mark the entry-point of the neural network, we use a `PipeOpTorchIngress`, for which three different flavors exist:

* `po("torch_ingress_num")` for numeric data
* `po("torch_ingress_categ")` for categorical columns
* `po("torch_ingress_ltnsr")` for `lazy_tensor`s

Because the imagenet task contains only one feature of type `lazy_tensor`, we go for the last option:

```{r}
architecture = po("torch_ingress_ltnsr")
```

We now define a relatively simple convolutional neural network.
Note that in the code below `po("nn_relu_1")` is equivalent to `po("nn_relu", id = "nn_linear_1")`.
This is needed, because `mlr3pipelines::Graph`s require that each `PipeOp` has a unique ID.

What we can further notice is that we don't have to specify the input dimension for the convolutional layers, which are inferred from the task during `$train()`ing.
This means that our `Learner` can be applied to tasks with different image sizes, each time building up the correct network structure.

```{r}
architecture = architecture %>>%
  po("nn_conv2d_1", out_channels = 64, kernel_size = 11, stride = 4, padding = 2) %>>%
  po("nn_relu_1", inplace = TRUE) %>>%
  po("nn_max_pool2d_1", kernel_size = 3, stride = 2) %>>%
  po("nn_conv2d_2", out_channels = 192, kernel_size = 5, padding = 2) %>>%
  po("nn_relu_2", inplace = TRUE) %>>%
  po("nn_max_pool2d_2", kernel_size = 3, stride = 2)
```

We can now continue with specifying the classification part of the network, which is a dense network that repeats a layer twice:

```{r}
dense_layer = po("nn_dropout") %>>%
  po("nn_linear", out_features = 4096) %>>%
  po("nn_relu_6")
```

In order to repeat a segment from a network multiple times, we can use `po("nn_block")`, which we here repeat twice.
Then, we follow with the output head of the network, where we don't have to specify the number of classes, as they can also be inferred from the task

```{r}
classifier = po("nn_block", dense_layer, n_blocks = 2L) %>>%
  po("nn_head")
```

Next, we can combine the convolutional part with the dense head:

```{r}
architecture = architecture %>>%
  po("nn_flatten") %>>%
  classifier
```

Below, we display the network:

```{r}
architecture$plot(html = TRUE)
```

To turn this network architecture into an `mlr3::Learner` what is left to do is to configure the loss, optimizer, callbacks, and training arguments, which we do now:
We use the standard cross-entropy loss, SGD as the optimizer and checkpoint our model every 20 epochs.

```{r}
checkpoint = tempfile()
architecture = architecture %>>%
  po("torch_loss", t_loss("cross_entropy")) %>>%
  po("torch_optimizer", t_opt("sgd", lr=0.01)) %>>%
  po("torch_callbacks", 
    t_clbk("checkpoint", freq = 20, path = checkpoint)) %>>%
  po("torch_model_classif",
    batch_size = 32, epochs = 100L, device = "cuda")

cnn = as_learner(architecture)
cnn$id = "cnn"
```

This created `Learner` now exposes all configuration options of the individual `PipeOp`s in its `$param_set`, from which we show only a subset for readability:

```{r}
as.data.table(cnn$param_set)[c(32, 34, 42), 1:4]
```

We can still change them, or if we wanted to, even tune them!
Below, we increase the number of blocks and latent dimension of the dense part, as well as change the learning rate of the SGD optimizer.

```{r}
cnn$param_set$set_values(
  nn_block.n_blocks = 4L,
  nn_block.nn_linear.out_features = 4096 * 2,
  torch_optimizer.lr = 0.2
)
```

```{r, include = FALSE}
cnn$param_set$set_values(
  torch_model_classif.epochs = 1L,
  torch_model_classif.device = "cpu"
)
imagenet$filter(1L)
```

Finally, we train the learner on the task:

```{r, include = FALSE}
# otherwise downloading might show in the code-snippet below
imagenet$data()
```


```{r}
cnn$train(imagenet)
```
