---
title: "Defining an Architecture"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Defining an Architecture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we will show how to build neural network architectures as `mlr3pipelines::Graphs`s.
We will create a simple CNN for the tiny-imagenet task, which is a subst of well-known Imagenet benchmark.

```{r, message = FALSE}
library(mlr3torch)
imagenet = tsk("tiny_imagenet")
imagenet
```

The central ingredients for creating such graphs are `PipeOpTorch` operators.

To mark the entry-point of the neural network, we use a `PipeOpTorchIngress`, for which three different flavors exist:
* `po("torch_ingress_num")` for numeric data
* `po("torch_ingress_categ")` for categoricalc columns
* `po("torch_ingress_ltnsr")` for `lazy_tensor`s

Because the imagenet task contains only one feature of type `lazy_tensor`, the latter is the correct choice:

```{r}
architecture = po("torch_ingress_ltnsr")
```

We now define a relatively simple convolutional neural network.
Note that `po("nn_relu_1")` is equivalent to `po("nn_relu", id = "nn_linear_1")`.
This is needed, because `mlr3pipelines::Graph`s require that each `PipeOp` has a unique ID.
What we can notice that we don't have to specify the input dimension for the convolutional layers, which are inferred from the task during `$train()`ing.
This means that our `Learner` can be applied to tasks with different image sizes.

```{r}
architecture = architecture %>>%
  po("nn_conv2d_1", out_channels = 64, kernel_size = 11, stride = 4, padding = 2) %>>%
  po("nn_relu_1", inplace = TRUE) %>>%
  po("nn_max_pool2d_1", kernel_size = 3, stride = 2) %>>%
  po("nn_conv2d_2", out_channels = 192, kernel_size = 5, padding = 2) %>>%
  po("nn_relu_2", inplace = TRUE) %>>%
  po("nn_max_pool2d_2", kernel_size = 3, stride = 2)
```

We can now continue with specifying the classification part of the network, which is a dense network that repeats a layer twice:

```{r}
dense_layer = po("nn_dropout") %>>%
  po("nn_linear", out_features = 4096) %>>%
  po("nn_relu_6")
```

In order to repeat a segment from a network multiple times, we can use `po("nn_block")`, which we here repeat twice.
Then, we follow with the output head of the network, where we don't have to specify the number of classes, as they can also be inferred from the task

```{r}
classifier = po("nn_block", dense_layer, n_blocks = 2L) %>>%
  po("nn_head")
```

Next, we can combine the convolutional part with the dense head:

```{r}
architecture = architecture %>>%
  po("nn_flatten") %>>%
  classifier
```

Below, we display the network:

```{r}
architecture$plot(html = TRUE)
```

To turn this network architecture into an `mlr3::Learner` what is left to do is to configure the loss, optimizer, callbacks, and training arguments, which we do now:
We use the standard cross-entropy loss, SGD as the optimizer and checkpoint our model every 20 epochs.
F

```{r}
checkpoint = tempfile()
architecture = architecture %>>%
  po("torch_loss", t_loss("cross_entropy")) %>>%
  po("torch_optimizer", t_opt("sgd")) %>>%
  po("torch_callbacks", 
    t_clbk("checkpoint", freq = 20, path = checkpoint)) %>>%
  po("torch_model_classif",
    batch_size = 32, epochs = 100L, device = "cuda")

cnn = as_learner(architecture)
cnn$id = "cnn"
```

This created `Learner` now exposes all configuration options of the individual `PipeOp`s in its `$param_set`, from which we show only a subset for readability:

```{r}
as.data.table(cnn$param_set)[c(32, 34, 42), 1:4]
```

We can still change them, or if we wanted to, even tune them!
Below, we increase the number of blocks and latent dimension of the dense part, as well as change the learning rate of the SGD optimizer.

```{r}
cnn$param_set$set_values(
  nn_block.n_blocks = 4L,
  nn_block.nn_linear.out_features = 4096 * 2,
  torch_optimizer.lr = 0.2
)
```

```{r, include = FALSE}
cnn$param_set$set_values(
  torch_model_classif.epochs = 1L,
  torch_model_classif.device = "cpu"
)
imagenet$filter(1L)
```

Finally, we train the learner on the task:


```{r}
cnn$train(imagenet)
```
